---
title: Introduction {#sec-introduction}
number-sections: true
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringdist)
library(ggdist)
library(patchwork)

# set ggplot theme and colour palette
my_theme <- theme_minimal() +
	theme(panel.grid = element_blank(),
		  axis.line = element_line(colour = "black"),
		  text = element_text(size = 12, colour = "black"),
		  axis.text = element_text(colour = "black"))

theme_set(my_theme)

clrs <- c("#AA0E28", "#ED1D3F", "#F47B8F", "#70B3FF", "#004CA3", "#002147")

options(
	ggplot2.ordinal.fill = clrs[c(1, 4, 5)],
	ggplot2.ordinal.colour = clrs[c(1, 4, 5)],
	ggplot2.discrete.fill = clrs[c(1, 4, 5)],
	ggplot2.discrete.colour = clrs[c(1, 4, 5)],
	ggplot2.continuous.fill = ggplot2::scale_color_gradient,
	ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

set.seed(1234)
```

Most caregivers would be surprised to find that infants show signs of language comprehension as early as six months of age. There are indeed reasons to be surprised. The linguistic input presents infants with a fairly arbitrary set of sounds and visual cues, and the child faces the tasks of identifying which of them correspond to relevant linguistic units in their native language, and associating them to their corresponding referents. This task is far from trivial, yet infants succeed early in age, laying the foundations of am initial *mental lexicon*. In the present dissertation, we delve into early word acquisition through the lens of a particular case of language learner: the bilingual infant.

Bilingual infants are exposed to a dual linguistic input, which requires a number of computations that monolingual infants do not face. For instance, bilinguals learn two sets of words, one in each language, for the same set of concepts. An English-Spanish bilingual infant learns the words *cat* and *gato* for the same referent. Despite the increased complexity of their linguistic input, bilinguals do not fall behind their monolingual counterparts in their trajectories of vocabulary growth. The particular mechanisms that allow bilinguals to keep up with monolinguals are still unclear. In this dissertation, we investigate the possibility that bilinguals exploit *cognateness* (i.e., the phonological similarity between translation equivalents) to facilitate the acquisition of new words.

We consider the bilingual case as a natural extension of the monolingual case, which may provide insights about the early lexicon in both populations. For this reason, this introductory chapter begins with a brief summary of monolingual language acquisition during the first year of life, from early linguistic experience, to the shift from language-universal to language-specific perception abilities, and the formation of the first phonological representations of familiar word-forms. We then characterise the case of bilinguals, with an emphasis on the relationship between the dual linguistic exposure they receive, and their trajectories of lexical acquisition. We point out some unresolved questions concerning the impact of linguistic similarity on vocabulary growth, and recent suggestions of a facilitatory role of cognateness. Next, we discuss the *parallel activation* hypothesis of such facilitation effect, contextualised within the language non-selective account of the bilingual lexicon. Finally, we summarise the available evidence of language non-selectivity in the initial lexicon, and the methodological issues therein encountered. We conclude this chapter describing the aims and structure of the present dissertation.

## The foundations of a lexicon

### From speech sounds to lexical representations

Infants' earliest demonstrations of language familiarity trace back to their last weeks of pre-natal life. Hearing foetuses are exposed to auditory input from the 20th week of gestation [@eggermont2011morphological], and start reacting to sounds between 24 and 36 weeks [@dipietro2013physiological]. From 36 weeks on, foetuses notice when the gestational parent is talking [@voegtline2013nearterm; @kisilevsky2009fetal; @decasper1980human], and can identify familiar melodies [@decasper1994fetal]. At birth, infants' perceptual abilities are already tuned to pre-natal experience with language. Newborns can discriminate between linguistic and non-linguistic sounds [@vouloumanos2004tuned; @ecklund-flores1996asymmetric; @may2018specificity], and between their native language and non-native languages [@mehler1988precursor; @moon1993twodayolds]. These early language discrimination abilities are shaped by infants' familiarity with some suprasegmental properties of the native language, namely prosody [@gervain2018role]. For instance, during the first two months of life, language discrimination is restricted to pairs of languages from different rhythm classes [@nazzi1998language; @ramus1999correlates; @abboub2016prosodic; @ramus2000language; @byers-heinlein2010roots; @pena2003sounds; @mehler1988precursor; @cooper1990preference]. It is not until six months of age that infants use segmental information to discriminate between their native language and others. Before six months of age, infants discriminate between virtually any pair of phonemes [@bertoncini1987discrimination; @eimas1971speech; @bertoncini1987discrimination; @aslin1981discrimination]. Between six and 12 months of age [but see @zacharaki2022perceptual for evidence at earlier ages], infants' perceptual abilities start tuning to the phoneme repertoire of their native language. By their first birthday, infants can only perceive phonemic contrasts present in their native language [@werker1984crosslanguage; @werker1983developmental; @kuhl1991human; @best1994emergence; @kuhl2006infants; @segal2016effect].

As infants tune their speech perception abilities to the phoneme repertoire of their native language, they also start storing their first representations of familiar word-forms^[Attempting to provide a clear definition of *word* goes beyond the aims of the present dissertation. We adhere to @jackendoff2002complexity's broad definition of *word* as a "finite list of structural elements that are available to be combined through a finite set of combinatorial principles, or *grammar*."]. This task is not trivial, as young infants are rarely exposed to single-word utterances. Instead, infants' speech input mostly consists of utterances and other multi-word combinations, providing isolated words in rare occasions [e.g., @aslin1996models]. To identify word boundaries, infants rely on multiple mechanisms and cues available in the speech signal, like phonotactics [@jusczyk1994infants; @friederici1993phonotactic; @friedrich2005phonotactic], phrase- and sentence-level prosodic contours [@soderstrom2003prosodic; @christophe1996bootstrapping; @gout2004phonological; @jusczyk1992perception], the predominant stress patterns of their native language [@thiessen2007learning; @cutler1990exploiting; @cutler1988role], or the statistical co-occurrence between syllables [@saffran1996statistical; @goodsitt1993perceptual; @saffran2001words]. By the end of their first year, infants recognise familiar word-forms [@vihman2004role; @halle1996format; @goodsitt1993perceptual], even if embedded in continuous speech [@jusczyk1995infants]. These initial phonological representations are encoded in great phonetic detail: when infants are presented with subtle mispronunciations of known word-forms (e.g., one phonetic feature apart), they struggle to recognise them as familiar [@swingley200511montholds; @mani2007phonological; @ballem2005phonological; @tamasi2017pupillometry]. Being able to parse the speech stream into familiar word-forms is a critical milestone in early language acquisition: phonological representations of word-forms lay the foundations of an early lexicon.

```{r wordbank}
wb <- here::here("data", "01-introduction", "wordbank_administration_data.csv") |> 
	arrow::read_csv_arrow() |>
	add_count(language) |>
	filter(typically_developing,
		   monolingual,
		   language!="English (British)",
		   between(age, 10, 30))
```

The foundations of the lexicon are laid by the first associations between phonological representations of familiar word-forms and their referents^[Although the role of the grammar in the early lexicon falls outside of the scope of the present dissertation, we acknowledge that grammatical information is incorporated in the lexicon, together with concepts and word-forms [@caramazza1997how; @krauska2023moving], but we do not adhere to any particular account of way in which grammatical information may be encoded in the lexicon.]. Evidence from inter-modal paradigms of word recognition [@hirsh-pasek1996intermodal; @delleluche2015methodological] shows that, by six months of age, infants are already forming their first lexical representations. For instance, Bergelson and Swingley [-@bergelson2012months; -@bergelson2015early] presented pairs of familiar pictures to infants at different age points, from six to 20 months. In each trial, their caregiver named one of the pictures, participants' looking preference for the named picture was measured, and it was interpreted as a proxy of word recognition. Overall, the authors found evidence of target looking preference in all age groups, suggesting that word comprehension emerges from six months of age. This initial lexicon is sparse: according to parental reports of word acquisition in many languages, most infants barely understand more than 30 words before their first birthday [e.g., @bates1994developmental; @jackson-maldonado1993early; @kern2007lexicon; @fenson1994variability]. Early comprehension is usually associated with words of high lexical frequency, which refer mostly to people, interjections, body parts, colours and food [@tardif2008baby; @parise2012electrophysiological; @tincoff2012sixmontholds; @tincoff1999beginnings; @frank2021variability; @forbes2019infants; @campbell2022scope; @marchman2002concurrent].

### Vocabulary growth

During the second year of life, this initial lexicon undergoes rapid development, in quantitative and qualitative terms. Evidence for this is mostly provided by parental reports of vocabulary size across ages, collected using the MacArthur-Bates Communicative Development Inventory (MCDI) [@bates1994developmental; @fenson1994variability], or its adaptations to other languages and populations. This questionnaire includes a vocabulary checklist, which contains a list of words---generally between 400 and 600 words---to which caregivers respond if their child *Understands*, *Understand and Says*, or if the child does not understand or say the word. The response options may vary between versions and adaptations of the questionnaire, but the end product is generally the same: an estimate of participants' receptive (number of words the child *understands*) or productive (number of words the child *says*) vocabulary size, obtained by aggregating caregivers' responses across words in the vocabulary checklist. Despite the apparent subjectivity of parents' answers, estimates of vocabulary size produced by this questionnaire show moderate to high intra-participant reliability [@reese2000predictive; @feldman2005concurrent; @rescorla2005concurrent; @frank2021variability] and evidence of concurrent and predictive validity (e.g., in experimental settings, participants are more likely to successfully recognise auditory words which their caregivers have marked as acquired) [@jahn-samilo2001vocabulary; @pan2004measuring; @can2013longterm; @dale1991validity; @werker2002infants; @swingley2000spoken; @bates2013emergence; @robinson1999comparing].

@fig-wordbank illustrates how rapidly children's vocabulary grows during their second year of age. This figure shows receptive and productive vocabulary size norms available in the Wordbank database [@frank2017wordbank], which contains responses to the CDI by families of children learning many different languages across the world. We show vocabulary size norms for `r format(n_distinct(wb$child_id), big.mark = ",")` monolingual children learning `r n_distinct(wb$language)` distinct languages. As it can be seen in the figure, at 18 months of age, the average infant understands around 81 words and produces 42 words. At 24 months, they know around 243 words and produce 235 words. By 30 months, they understand around 507 words and produce 501 words. This rapid vocabulary growth, sometimes referred to as the *vocabulary spurt* [@goldfield1990early; @bloom2002how], impacts infants' trajectories of word recognition [@fernald2013ses; @fernald1998rapid; @marchman2008speed]. @fernald1998rapid examined the speed at which 15-to-24-month-old infants looked at named target pictures, compared to distractor pictures. Although infants of all ages showed a looking preference for the target pictures---suggesting successful spoken word recognition---older infants did so faster. At 24 months, infants started directing their gaze towards the target picture before the offset of the spoken label [see also @swingley1999continuous]. At 15-months, infants shifted their gaze only after the end of the spoken label. Overall, these results suggest an improvement in the efficiency of spoken word recognition during the second year of life. The mechanisms behind this accelerated lexical acquisition and processing are still under debate [@bergelson2020comprehension; @mcmurray2007defusing; @tomasello2000socialpragmatic; @bloom2002how].

```{r fig-wordbank}
#| label: fig-wordbank
#| fig-cap: "Wordbank norms for receptive (A) and productive (B) vocabulary size in monolingual children. Vocabulary sizes are expressed as the number of words that caregivers reported to be acquired by their child, either as *Understands* (receptive vocabulary) or *Undertands and Says* (productive vocabulary). Data were collapsed for each month of age. Median vocabulary sizes for each age group are indicated with white points and lines. Intervals of different width (depicted with shades of colour) contain different porportions of the sample."
#| fig-width: 6
#| fig-height: 8
wb_fig <- wb |>
	pivot_longer(c(comprehension, production),
				 names_to = "type",
				 values_to = "size",
				 names_transform = tools::toTitleCase)

wb_summary <- wb_fig |>
	summarise(size = median(size),
			  .by = c(type, age))

plot_comp <- wb_fig |>
	filter(type=="Comprehension") |>
	ggplot(aes(age, size)) +
	facet_wrap(~type) +
	stat_interval(.width = c(0.95, 0.89, 0.75, 0.67, 0.50),
				  size = 7.5,
				  position = "dodge") +
	geom_line(data = filter(wb_summary, type=="Comprehension"),
			  colour = "white",
			  linewidth = 1) +
	geom_point(data = filter(wb_summary, type=="Comprehension"),
			   colour = "white",
			   size = 2.25) +
	scale_colour_manual(values = colorRampPalette(c("grey95", clrs[6]))(5))


plot_prod <- wb_fig |>
	filter(type=="Production") |>
	ggplot(aes(age, size)) +
	facet_wrap(~type) +
	stat_interval(.width = c(0.95, 0.89, 0.75, 0.67, 0.50),
				  size = 7.5,
				  position = "dodge") +
	geom_line(data = filter(wb_summary, type=="Production"),
			  colour = "white",
			  linewidth = 1) +
	geom_point(data = filter(wb_summary, type=="Production"),
			   colour = "white",
			   size = 2.25) +
	scale_colour_manual(values = colorRampPalette(c("grey95", clrs[1]))(5))


plot_comp + plot_prod +
	plot_layout(ncol = 1) &
	plot_annotation(tag_levels = "A") &
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Proportion of the sample",
		 colour_ramp = "Proportion of the sample") &
	scale_x_continuous(breaks = seq(0, 30, 2)) &
	scale_y_continuous(breaks = seq(0, 1000, 100)) &
	theme(legend.position = c(0, 1),
		  legend.justification = c(0, 1),
		  panel.grid.major.y = element_line(
		  	colour = "grey",
		  	linetype = "dotted"))
```

### Dynamics of early lexical access and selection

There is some consensus on that the initial lexicon is only different from the adult lexicon in quantitative terms. Many discontinuities and non-linearities in the performance of infants and adults in word recognition tasks can be explained as the result of differences in the size of the lexicon. For instance, evidence of *cascaded activation* in the developing lexicon emerges at around 21 months, associated with changes in the size of the lexicon. Cascaded accounts of lexical processing describe how activation spreads across representations at different levels in the lexicon, as the speech signal unfolds, and is one of the most defining features of the adult lexicon [@dell1986spreadingactivation; @levelt1989language]. Most of these accounts assume, at least, three levels of representation: phonological, lexical, and semantic [@caramazza1997how]. Lexical representations embody associations between how words sound (*phonological representations*, or word-forms) and what words mean (*semantic representations*, or concepts). During word production and comprehension, representations from the three levels are activated (lexical access), and one of them is selected for comprehension or production (lexical selection). Which representations are activated across the three levels is determined by bottom-up and top-down sources of information. @fig-lexicon-mon illustrates the dynamics in word comprehension.

![Lexical access and selection in the monolingual lexicon. The speech signal produced by a speaker uttering the Catalan word *COTXE* [car] activates phonological segments in the listener's mental lexicon. Activation then spreads across lexical representations containing such segments. In the illustration, a cohort of words also starting with phoneme /k/ is activated. The phonological form associated with the word *COTXE* receives the strongest activation, and is ultimately selected. In non-selected, but accessed, lexical representations, activation spreads also to the semantic layer, resulting in the activation of non-selected semantic representations. Note that, for simplicity, we have used phonemes to depict phonological units, but we acknowledge that phonological representations may be encoded at the lower level of phonetic features. ](../_assets/img/lexicon-mon.png){#fig-lexicon-mon}

Spoken word recognition starts with the acoustic-phonetic processing of the speech signal. Phonological representations (e.g., articulatory features, syllables, stress patterns) in the listener's mental lexicon are activated according to how well they match the phonetic features and co-articulatory information in the speech stream. Lexical representations whose associated phonology matches the input are activated. The set of lexical candidates then is modulated by multiple factors: the degree of (mis)match with the unfolding acoustic input, the dynamics of competition between lexical candidates [@luce1998recognizing; @mcclelland1986trace; @norris1994shortlist], and top-down information like word frequency [@kukona2011time], or grammatical, suprasegmental, and semantic constraints provided by the sentential context [@marslen-wilson1988lexical; @tyler1983quantifying; @grosjean1980spoken]. Finally, the best-matching lexical representation is selected, and its associated meaning is activated for comprehension. This sequence of events occurs in a *cascaded* fashion. Activation spreads within and percolates across levels of representation while lexical selection is still ongoing. Lexical representations sharing phonological similarity [e.g., @marslen-wilson1987functional; @marslen-wilson1978processing; @grosjean1980spoken; @mcclelland1986trace; @luce1995similarity] or semantic features [@neely1977semantic; @collins1975spreadingactivation; @mirman2009dynamics] are co-activated, and semantic representations of activated lexical units are in turn retrieved, even if their associated lexical representation is ultimately not selected.

A clear proof of cascaded activation at the phonological and semantic levels was provided by @allopenna1998tracking. The authors designed a word-recognition task in which (adult) participants were presented with four pictures on screen in each trial. Participants would then listen to a command sentence like "Pick up the *casket*; now put it below the diamond", and perform the action. The authors manipulated the phonological relationship between the target object's label (e.g., *casket*), and the label of each of the three distractors. Two of the distractors were phonologically related to the target, sharing onset (e.g., *castle*) or offset (e.g., *basket*), respectively. The other distractor was phonologically unrelated to the target (e.g., *nickel*). After the onset of the auditory target label, phonological distractors attracted participants' eye fixations more than unrelated distractors. Distractors sharing phonological onset elicited the effect at an earlier time window than those sharing phonological offset. Such time course of distractor fixations suggests that the auditory presentation of the target label activated not only its (subsequently selected) phonological representation in participants' lexicon, but also that of phonologically related representations. In summary, paradigms of spoken word recognition have proven useful in the investigation of the structure of the adult lexicon: phonological and semantic links between lexical representations are revealed as the phonological or semantic relationship between the stimuli is manipulated.

Priming paradigms of spoken word recognition have also provided strong support for the existence of phonological and semantic links between lexical representations before infants' second birthday. @arias-trejo2009lexical found that infants' spoken word recognition (e.g., *dog*) was interfered by the previous presentation of a semantically related word (e.g., *cat*). This effect was found in 21-month-olds, but not in 18 month-olds. @styles2009how reported semantic priming in 24-month-old participants but not in 18 month-old participants. Earlier evidence of semantic priming has been provided by electrophysiological methods [e.g., @rama2013development; @friedrich2005lexical]. Even in the absence of visual referents, 24-month-old infants can tell when the words embedded in the speech stream are semantically related [@willits2013toddlers]. Overall, effects of semantic relations in word recognition are observable in the initial lexicon.

Phonological priming effects also emerge from 18-months of age. Mani and Plunkett [-@mani2010infant; -@mani2011phonological] developed an implicit naming task in which each trial started with the *silent* presentation of a prime picture. Then, a target-distractor picture pair was presented side-by-side. Finally, the target picture's label was presented auditorily. The authors manipulated the phonological overlap between the prime label and the target label, so that in half of the trials both labels were phonologically related, sharing phonological onset (*cat*-*cup*), or phonologically unrelated (*ball*-*comb*). Prime and target-distractor pairs were semantically unrelated. At 18 and 21 months of age, English monolingual infants showed different target looking preference patterns after phonologically related prime pictures, compared to after phonologically unrelated primes. This result suggests that infants implicitly named prime pictures---despite such pictures were presented in silence---and that the phonology of the resulting labels interacted with the subsequent auditory target recognition. The direction of the effect was different in both age groups. At 18 months, infants showed stronger target preferences after being previously presented with a phonologically related prime. But 21-month-olds' target preference was weaker after phonologically related primes, compared to after phonologically unrelated primes.

@mani2011phonological interpreted this shift from facilitation to inhibition as the consequence of the increase in vocabulary sizes of the older infants. The authors suggested that, in sparser lexicons, the fewer number of phonologically related lexical representations might have led to faster selection of the target word. In larger lexicons, lexical selection would be delayed by the cascaded activation of a larger number of phonologically related lexical representations. In support of this hypothesis, the authors found a positive significant correlation between the size of the interference effect in 21-month-old infants with two critical variables: participants' vocabulary size, and the size of the phonological cohort of the presented words. Results by @avila-varela2021longitudinal provided converging evidence with German monolinguals: larger interference effects were found in participants with larger vocabulary sizes, even after controlling for age.

@chow2017spokenword provided further evidence of the emergence of semantic *and* phonological links, and its association with lexical growth. The authors adapted the spoken word recognition paradigm by @huettig2007tug to explore 24-to-30-month-old toddlers. In each trial, participants were presented with four pictures. Four seconds after, a word was auditorily presented. The authors analysed participants' visual fixations to each picture after the auditory label was presented. In some trials, the auditory label referred to one of the pictures. In other trials, it did not refer to any of the pictures. In trials in which none of the pictures corresponded to the presented label, one of the pictures was a phonological distractor (its label shared phonological onset with the presented label). Another picture was a semantic distractor (its referent belonged to the same taxonomic category as the referent of the presented word). For instance, participants were presented with the pictures of a sandwich, a bus, a cat, and a dress. Then they would hear the carrier phrase "Look at the *bee*!". The picture of the bus would be a phonological distractor (both words start with /\textipa{b}/), and the cat would be a semantic distractor (bees and cats are animals). Participants' visual fixations revealed a preference for the phonological distractor at earlier stages of the post-naming phase, and a preference for the semantic distractor at later stages of the trial. These results are suggestive of a cascaded effect in spoken word recognition in toddlers, in which the presentation of the auditory word activated phonologically and semantically related lexical representations. Visual preference for the distractors was stronger in participants with larger vocabulary sizes, pointing to an association between the strength of the connections across related lexical representations and the growth of the lexicon.

In summary, previous studies on lexical access in monolingual toddlers support a cascaded activation account of lexical access during the first stages of lexical development, and point to a continuity between the initial lexicon and the adult lexicon, bridged through vocabulary growth during the second year of life.

## Building a lexicon in two languages

### Defining bilingualism

One may consider *bilingualism* a quite ill-defined term. In some studies participants are classified as bilinguals with very low exposure and command of a second language, while other studies require perfect, native-monolingual proficiency of both languages for a participant to be classified as bilingual. Since the present dissertation concerns simultaneous bilingualism in infancy, we identify bilingualism with dual language exposure. Quantitative language input is conventionally operationalised as the cumulative amount of time the infant has spent interacting with people who speak to them on a regular basis (e.g., parents, grandparents, siblings, daycare). In bilinguals, this linguistic input is divided into two languages, which may be spoken by the same or different people, at the same time or at different times.

Several instruments are available to measure language exposure in bilinguals, usually involving a semi-structured interview with the caregivers or detailed questionnaires that caregivers complete [e.g., @bosch2001evidence; @cattani2014how; @deanda2016language; @orena2020what; @byers-heinlein2020maple]. The end-product is generally a proportion of exposure to each of the languages, known as *Degree of Exposure* (DoE). The DoE reflects the cumulative amount of input that the child has received in one language, relative to the other. Although this measure is subject to biases and inaccuracies induced by the subjective judgement of caregivers, there is evidence of its external validity, as suggested by its high correlation with direct measures of infant-directed speech [@orena2020what]. The underlying assumption behind this measurement approach is that a child's linguistic experience can be quantified into a continuous measure of bilingualism. This score ranges from 0% exposure to a second language (monolinguals) to 50% (perfectly balanced exposure to both languages). The language exceeding 50% DoE is referred to as the *dominant* language of the child, and the other language is referred to as the *non-dominant* language. For methodological simplicity, some studies classify monolinguals and bilinguals into discrete categories. Although there is not a strict threshold to consider a child as bilingual, most studies consider a child as bilingual if they are exposed to a second language at least 20% of the time [@byers-heinlein2015methods; @byers-heinlein2021multilab; @rocha-hidalgo2023defining] (see @fig-bilingualism for a visual illustration). This is the cut-off adopted in the present dissertation, whenever such classification into discrete groups is methodologically sound.

![Visual illustration of the classification of participants into monolinguals and bilinguals in the present dissertation. Participants' relative exposure to Catalan and Spanish ranged from completely Catalan (in blue) to completely Spanish (in red). In agreement with the consensus in the field, we classified participants into monolinguals and bilinguals by setting a threshold of 20% degree exposure to the language of least exposure. Participants with a relative exposure to a second language higher than 20% were classified as bilinguals, and as monolinguals otherwise.](../_assets/img/bilingualism.png){#fig-bilingualism}

### Bilingual word acquisition

Bilinguals acquire language in a more challenging set of circumstances than monolinguals. First, there is little reason to think that bilinguals receive a larger amount of language exposure than their monolingual peers. Given that bilinguals’ linguistic input is divided into two languages, they receive less input in each of their languages than monolinguals do. Second, bilinguals need to learn two codes, which might overlap to different degrees: two phonologies, two grammatical systems, and two lexicons. Third, the referential context in which bilinguals build a lexicon is also more complex. Bilinguals frequently learn at least two labels for the same referent, one for each language. The concepts behind both labels may not align perfectly. For instance, a child learning English and Spanish might learn the words *finger* and *dedo* (its Spanish translation). While both words might be used to refer to the same referent (e.g., the index *finger*), the word finger refers to eight appendices (four in each hand), whereas *dedos* refers to 20 appendices (five in each hand and five in each foot). In summary, bilinguals are presented with a more complex linguistic environment, facing additional demands compared to their monolingual counterparts.

Such increased cognitive demands do not keep bilinguals from reaching major language acquisition milestones at similar ages as monolinguals [see @sebastian-galles2020bilingual for review], as observed in language discrimination [@bosch1997nativelanguage; @bosch2001evidence; @byers-heinlein2010roots; @weikum2007visual; @sebastian-galles2012bilingual; @nacargarcia2018evoked], discrimination of native phonemic contrasts [@bosch2003simultaneous; @albareda-castellot2011acquisition; @sundara2006languageexperience; @burns2007development; @sundara2008development], and word-form segmentation [@nazzi2006early; @bosch2013rapid; @antovich2018learning; @houston2000role; @tsui2021dual; @polka2002role; @hurtado2014relative]. During the second year of life, bilinguals also undergo the same increase in vocabulary size as monolinguals [e.g., @pearson1994patterns; @hoff2012dual; @core2013total]. As in monolinguals, spoken word recognition becomes more efficient as the infant's lexicon expands [@deanda2020lexicalsemantic; @deanda2018lexical; @legacy2018vocabulary; @poulin-dubois2013lexical; @poulin-dubois2017translation]. Such changes are modulated by the relative amount of input that the bilingual infant receives from each language. For instance, bilinguals acquire words faster, and become more efficient at spoken word recognition in their dominant language, compared to the non-dominant language [@thordardottir2011relationship; @cattani2014how; @hoff2003specificity; @shneidman2013what; @floccia2018introduction; @marchman2010how]. Overall, the monolingual lexicon and the bilingual lexicon grow at a similar rate, but some studies suggest that bilingualism may modulate word learning trajectories.

@hoff2012dual collected vocabulary size data from a cohort of 103 infants at 22 and 30 months. Infants were being raised in English monolingual or English-Spanish bilingual homes in South Florida (United States). Both groups had equivalent socio-demographic backgrounds. Families of English monolingual infants filled in the MCDI, and families of English-Spanish bilinguals filled the MCDI and its Spanish adaptation, the *Inventario del Desarrollo de Habilidades Comunicativas* [@jackson-maldonado1993early]. Monolinguals showed the largest English vocabulary sizes, followed by English-dominant bilinguals, and Spanish-dominant bilinguals. The reverse pattern was observed for Spanish vocabulary sizes. When @hoff2012dual combined the vocabulary English and Spanish vocabulary sizes into a single measure of *total* vocabulary, monolinguals and bilinguals showed equivalent vocabulary sizes. These results indicate that bilinguals keep up with monolinguals in vocabulary growth [see also @pearson1993lexical; @pearson1994patterns; @poulin-dubois2013lexical; @core2013total; @byers-heinlein2023sometimes; @bosch2014first], and highlight the importance of comparing monolingual and bilingual trajectories of lexical acquisition sizes taking both languages into account [@bedore2005conceptual]. 

Not all bilingual populations have shown equivalent patterns of lexical development. For instance, studies in other populations of bilinguals have reported differences in total vocabulary size in monolinguals and bilinguals, and equivalent conceptual vocabulary sizes. This suggests that bilinguals may acquire a larger number of words across both languages than monolinguals, corresponding to an equivalent number of lexicalised concepts in both groups [@dehouwer2014bilingual; @siow2023double], or even higher [@kern2019lexical]. These results suggest that the specific pair of languages that bilinguals learn must be taken into consideration when investigating the early stages of lexical development. Participants in @hoff2012dual were learning English and Spanish. These two languages are relatively distant in typological terms. English is a Germanic language, while Spanish is a Romance language. In addition to their differences in grammar or phonology, English and Spanish also share fewer cognates than other pairs of languages like Italian and Spanish [@schepens2012distributions]. Recent studies have reported that bilinguals learning two languages that share a higher amount of cognates may acquire words at a faster rate than those learning two languages sharing fewer cognates [@floccia2018introduction; @gampe2021does]. This might explain some of the variability observed in the characterisation of lexical growth across different bilingual populations.

## Language non-selectivity in the bilingual lexicon

@floccia2018introduction conducted an impressive study in which CDI responses were collected from a large sample of 372 bilinguals at 24 months of age. These bilinguals were learning English and an additional language---one of Bengali, Cantonese, Dutch, French, German, Greek, Hindi-Urdu, Italian, Mandarin, Polish, Portuguese, Spanish, and Welsh. Participants were raised in the United Kingdom, and English was the dominant language for most of them. Using an edit distance-based metric of similarity, the authors calculated the phonological similarity between the translation equivalents present in the English vocabulary checklist, and in the vocabulary checklist of each of the other languages. The average phonological similarity between English and each pair of languages was taken as a proxy to their lexical similarity. Some language pairs had high lexical similarity (e.g., English-Dutch), while others had very little (e.g., English-Mandarin). The authors then estimated the association between lexical similarity and participants' language outcomes---measured as their receptive and productive vocabulary size in English and the additional language---while adjusting for the influence of other predictors like age, sex, socio-economic status, and several properties of participants' speech input. Bilinguals' vocabulary size in the additional language showed a lexical similarity effect, in which children learning two lexically similar languages showed larger vocabulary sizes in the additional language, while English vocabulary did not show such association with lexical similarity. Overall, the results suggest that infants may benefit from the lexical similarity between their languages--particularly, by the presence of cognates---and therefore bilinguals' trajectories of word acquisition should be considered in the context of the pair of languages being learned.

A facilitation effect at the lexical level would have important consequences for infants learning highly similar languages. For intance, the relative typological distance between Catalan and Spanish is very small, especially at the lexical level. Both are Romance languages, and share many cognates (i.e., form-similar translation equivalents). @fig-cat-spa-distance shows the overall lexical similarity between Catalan and Spanish, computed in the same way that @floccia2018introduction did. The lexical similarity between Catalan and Spanish is considerably larger than that between English and Dutch, the language pair with the highest similarity score in Floccia et al. This suggests that Catalan-Spanish bilinguals might be exposed to a substantially larger amount of cognates than bilinguals learning English and and additional language, like those in @floccia2018introduction and @hoff2012dual. If linguistic similarity, and especially lexical similarity, play a facilitative role in bilingual word acquisition, Catalan-Spanish bilinguals should show an even larger effect. This might explain why bilinguals in our database display larger vocabulary sizes than monolinguals, in contrast with the equivalent total vocabulary sizes reported by @hoff2012dual in English-Spanish bilinguals (learning two languages having less lexical similarity).

```{r fig-cat-spa-distance}
#| label: fig-cat-spa-distance
#| fig-cap: "Lexical similarity between Catalan and Spanish, and between English and each of the additional languages in Floccia et al. (2018). We first obtained the broad phonological transcriptions of the words included in the Catalan and Spanish vocabulary checklists of the BVQ. Catalan words were transcribed to the Central Catalan variant (the most prevalent in the Metropolitan Area of Barcelona), and Spanish words were transcribed to Castilian Spanish. Interjections and onomatopoeic words were excluded. We then computed the normalised Levenshtein between each pair of translation equivalents, and averaged them."
#| fig-height: 5
#| fig-width: 6
bvq_data <- readRDS(here::here("data", "01-introduction", "bvq.rds"))

flatten_xsampa <- function(x) {
	str_rm <- c("\\.", "\\\\", ",", "/", "?", "'", '"')
	str <- gsub(paste0(str_rm, collapse = "|"), "", x)
	str <- gsub("\\{", "\\\\{", str)
	return(str)
}

dist <- bvq_data$pool |>
	mutate(phon = flatten_xsampa(xsampa)) |>
	filter(!(semantic_category %in% c("Interjections"))) |>
	distinct(te, language, .keep_all = TRUE) |>
	pivot_wider(id_cols = c(te),
				names_from = language,
				values_from = phon,
				names_repair = janitor::make_clean_names) |>
	mutate(distance = stringdist::stringsim(catalan, spanish)) |>
	pull(distance) |>
	mean(na.rm = TRUE)

tibble::tribble(~language, ~dist,
				"Dutch", 0.2214,
				"Welsh", 0.2163,
				"German", 0.1975,
				"Italian", 0.1076,
				"French", 0.1034,
				"Bengali", 0.0941,
				"Hindi", 0.0899,
				"Spanish", 0.0874,
				"Polish", 0.0828,
				"Greek", 0.0807,
				"Portuguese", 0.0801,
				"Cantonese", 0.0422,
				"Mandarin", 0.0197) |>
	mutate(language = paste0("English-", language)) |>
	add_row(language = "Catalan-Spanish", dist = dist) |>
	mutate(study = if_else(grepl("English", language),
						   "Floccia et al. (2018)", "BVQ")) |>
	ggplot(aes(dist, reorder(language, dist),
			   fill = study)) +
	geom_col(colour = "white") +
	geom_text(aes(label = scales::percent(dist, accuracy = 0.01)),
			  size = 3,
			  hjust = 0,
			  position = position_nudge(x = 0.005)) +
	labs(x = "Lexical similarity",
		 y = "Language pair") +
	scale_fill_manual(values = clrs[c(2, 5)]) +
	scale_x_continuous(limits = c(0, 0.50),
		labels = scales::percent,
		breaks = seq(0, 1, 0.1)) +
	theme(legend.position = "none",
		  axis.title.y = element_blank(),
		  panel.grid.major.x = element_line(
		  	linetype = "dotted",
		  	colour = "grey"))

```

The facilitation effect of lexical similarity on vocabulary growth reported by @floccia2018introduction points to a possible cognateness facilitation effect on word acquisition, in which cognates are acquired at earlier ages than non-cognates. In this scenario, bilinguals learning languages that share more cognates would acquire words faster than those learning two languages that share fewer cognates. In line with this claim, previous studies had provided evidence in favour of an earlier age of acquisition for cognates. For instance, some studies have reported a larger proportion of cognates in bilinguals' lexicons than the proportion of non-cognates [@bosch2014first; @fabian2016investigating; @schelletter2002effect]. Other studies have found that bilinguals' performance in word recognition tasks is increased for cognate words, relative to non-cognate words [@gampe2018bilex]. More recent evidence in English-French bilinguals suggest that cognates may be more likely to be acquired than non-cognates at early ages [@mitchell2022cognates]. The specific mechanisms behind an earlier age of acquisition for cognates, however, are unclear. 

To explain the facilitation effect of lexical similarity on vocabulary growth, @floccia2018introduction suggested that the *parallel activation* of cognates during language exposure might boost their acquisition. The notion of *parallel activation* stems from the language non-selective hypothesis of bilingual lexical processing. This hypothesis states that bilinguals co-activate lexical representations from both languages during language production and comprehension, even in monolingual situations. This parallel co-activation is the result of cascaded activation spreading across the two languages at multiple levels of word recognition and production. This hypothesis was initially proposed to account for different results in adult bilingual research.

@marian1999activation presented Russian-English adult bilinguals with a cross-language interference task, divided in two blocks. In one block, participants would be presented with an instruction in English, like "Put the *marker* below the cross," after which two pictures were displayed on the screen. One depicted the target object (a stamp). The other depicted an object whose label in Russian shared phonetic features with the English target label (e.g., a stamp, *marka* in Russian). In the other block, the same procedure would be followed, but now with Russian as the target language. Participants were presented with the Russian instruction: "*Poloji* glaz *nije krestika*" [Put the *eye* below the cross]. The target picture would be an eye, and the distractor, a glove, whose English label shared phonetic features with the Russian target label. In both blocks, a control condition was run, in which target and distractor labels were phonetically unrelated. In line with @allopenna1998tracking's results in monolingual adults, after hearing the target label, bilinguals fixated the cross-language phonological distractor significantly more than the unrelated distractors. These findings suggest that participants activated phonologically related word-forms in both Russian *and* English, which affected their overt visual exploration patterns during the task. In this case, parallel activation is driven by bottom-up activation of words in both languages through phonology, at a sub-lexical level. Words that sound the same in both languages are activated, and enter the cohort during lexical selection. @fig-lexicon illustrates this sequence of events.

![Lexical access and selection in the language non-selective bilingual lexicon. The speech signal produced by a speaker uttering the Catalan word *cotxe* [car] activates phonological segments in the repertoire of a Catalan-Spanish bilingual listener. Activation spreads across lexical representations in both languages that contains such segments. In the illustration, a cohort of words that also start with the /k/ phoneme are activated. The word-form associated with *cotxe* receives the strongest activation, and is ultimately selected. In non-selected, but accessed, lexical representations, activation percolates to the semantic layer, resulting in the activation of non-selected semantic representations.](../_assets/img/lexicon.png){#fig-lexicon}

Parallel activation can also be the result of cascaded activation spreading across lexical representations in both languages, through their shared semantic features. Previous studies have provided strong evidence for parallel activation in exclusively monolingual tasks by comparing the processing of cognates and non-cognates. In @costa2000cognate, adult Catalan-Spanish bilinguals and Spanish monolinguals were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*). In the other half of the trials labels were non-cognates (*taula*-*mesa*, translations of *table*). Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Spanish monolinguals showed equivalent naming times for the two categories, as the distinction was meaningless to them. These results suggested that, in bilinguals, Catalan phonology was activated, facilitating Spanish word production: the visual recognition of the pictures led to the activation of their corresponding semantic representations, resulting in the cascaded activation of phonological representations in both languages.

In adults, the available evidence in favour of language non-selectivity in the bilingual lexicon is abundant across languages [@spivey1999cross; @bobb2020coactivation; @duyck2005translation; @hoshino2008cognate; @marian2003competing; @colome2001lexical; @costa2000cognate; @thierry2007brain; @dunabeitia2009masked; @costa1999lexical], including sign languages [@gimeno-martinez2021crosslinguistic; @giezen2016language; @vanheuven1998orthographic]. Modelling efforts have successfully implemented formal accounts of the bilingual lexicon by incorporating cross-language interactions [e.g., @kroll2010revised; @dijkstra2002architecture; @kroll1994category; @dijkstra2019multilink]. In summary, there is consensus about the fact that bilinguals constantly co-activate both languages during word comprehension and production, even in fully monolingual situations.

Previous studies have provided evidence for language non-selectivity in the initial bilingual lexicon [@bosma2020cognate; @jardak2019labels; @poulin-dubois2013lexical; @poulin-dubois2017translation; @singh2014one; @vonholzen2012language; @vonholzen2019impact; @poarch2012crosslanguage]. Priming paradigms of word recognition have been instrumental in this line of research. Using one of such paradigms, @vonholzen2012language reported evidence for cross-language phonological priming in 21- to 42-months-old children learning German and English. At the beginning of each trial, the authors presented an auditory prime word in English. Then, the auditory target label was presented in German. Finally the target and distractor pictures were presented side-by-side. Participants' looking preference towards the target was recorded, and interpreted as a proxy of target word recognition. The authors manipulated the phonological overlap between the prime and target labels. In some trials, the English prime label and the German target label labels were phonologically related through translation: the prime label did *not* overlap with the target label in English (e.g., *leg*-*Stein* [*stone*]), but its translation in German did [*Bein*]. In the rest of the trials, prime and target labels were phonologically unrelated in both languages. The authors found a stronger target picture looking preference in unrelated trials, compared to trials in the priming-through-translation condition. This suggests that participants activated the German translation of the auditory English prime word, and that this interfered with the recognition of the auditory German target word (which was phonologically similar to its German translation).

Priming studies in which words from both languages are presented during the same experimental session or even within the same trial [e.g., @vonholzen2012language; @floccia2020translation; @jardak2019labels; @singh2014one] present an important methodological pitfall. In these tasks, participants are introduced in a bilingual context, in which the overall degree of activation of lexical representations in both languages may be artificially increased [@grosjean1997bilingual]. This might have contributed, to some extent, to the strength of the parallel activation reported in these studies. It is possible that, in the daily life of a sizable amount of bilingual children, speech input in each language takes place at separate times and it is unlikely that speech input in each language takes place within the same communicative act. If this is the case, the practical relevance of language non-selectivity in vocabulary growth might be smaller than anticipated.

In summary, the parallel activation hypothesis suggested by @floccia2018introduction to explain the cognateness facilitation effect relies on the language non-selectivity of the early lexicon. But the evidence for the developing lexicon for such language non-selectivity builds on experimental paradigms that may overestimate the amount of co-activation between the two languages, as such paradigms put participants in a bilingual context that might forcibly lead to both languages being active.

## The present dissertation

The aim of this dissertation is to explore the impact of language non-selectivity on the developing bilingual lexicon, and its potential role in the facilitation effect of lexical similarity during vocabulary growth. In @sec-chapter2, we put forward an account of bilingual word acquisition, inspired by accumulator models of word language acquisition [@hidaka2013computational; @mcmurray2007defusing; @mollica2017how; @kachergis2022standard]. This model provides a mechanistic explanation for the facilitative effect of lexical similarity reported by @floccia2018introduction. It describes an interplay between the lexical frequency of a word, the child's relative exposure to each language, and cross-language phonological similarity. In this model, bilinguals accumulate experience with words in one language, even when listening to words in the other language. We contrast this model against vocabulary data from bilinguals, collected using an *ad hoc* online questionnaire. This questionnaire was inspired by the CDI, adapted to the population of bilinguals learning Catalan and Spanish in the Metropolitan Area of Barcelona (Spain). We modelled the acquisition trajectories of 302 translation equivalents, testing the role of cognateness and its interaction with relative language exposure and lexical frequency.

The predictions tested in @sec-chapter2 rely on the assumption that bilingual infants co-activate translation equivalents in both languages, even in monolingual situations. Although there is evidence in favour of cross-language activation in toddlers, as just said, previous studies have relied on experimental paradigms in which participants listen to words from both languages in each trial. This puts participants in a bilingual context in which parallel activation might result from an overall stronger activation of both languages induced by the experimental task, and not because of regular cascaded activation across languages. In @sec-chapter3, adapting the implicit naming paradigm by Mani and Plunkett [-@mani2010infant; -@mani2011phonological], we test parallel activation in an exclusively monolingual task. As described in the Methods section of @sec-chapter2, this paradigm is a word recognition task in which priming stimuli are pictures presented in silence. Previous studies have shown that infants at 18 months and older lexicalise pictures presented in silence [@duta2012erp; @styles2015infant; @mani2010infant; @mani2011phonological]. In particular, Mani and Plunkett [-@mani2010infant; -@mani2011phonological] found that word recognition is influenced by the prior presentation of a picture whose associated label is phonologically related to the target word. By manipulating the cognate status of the label associated with the prime image, we tested parallel activation in bilingual toddlers while avoiding the presentation of words in both languages.

The final chapter of this dissertation provides a summary of the findings of @sec-chapter2 and @sec-chapter3. We discuss their implications for the current understanding of the initial lexicon, and of how bilingualism shapes infants' trajectories of word acquisition.

