---
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects
targets::tar_config_set(store = here::here("_targets"),
						script = here::here("_targets.R"))

targets::tar_load_globals()

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials,
	  attrition_participants,
	  bvq_data,
	  gaze,
	  data_bcn,
	  data_oxf
	))

targets::tar_load(
	c(model_fits_bcn,
	  model_loos_bcn,
	  model_fits_oxf,
	  model_loos_oxf
	))

# prepare data
attrition <- lst(participants,
				 attrition_participants,
				 attrition_trials) |>
	reduce(inner_join, by = join_by(session_id)) |>
	unnest_wider(matches("all"), names_sep = "_") |> 
	rename_with(\(x) gsub("is_valid_gaze_all_|is_valid_vocab_all", "", x))

library(knitr)
library(kableExtra)
library(ggplot2)
library(gt)
library(gtExtras)
library(patchwork)
library(english)
library(gtExtras)

# set ggplot theme and colour palette
my_theme <- theme_minimal() +
	theme(panel.grid = element_blank(),
		  axis.line = element_line(colour = "black"),
		  text = element_text(size = 12, colour = "black"),
		  axis.text = element_text(colour = "black"))

theme_set(my_theme)

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(ggplot2.ordinal.fill = clrs[c(1, 4, 5)],
		ggplot2.ordinal.colour = clrs[c(1, 4, 5)],
		ggplot2.discrete.fill = clrs[c(1, 4, 5)],
		ggplot2.discrete.colour = clrs[c(1, 4, 5)],
		ggplot2.continuous.fill = ggplot2::scale_color_gradient,
		ggplot2.continuous.colour = ggplot2::scale_color_gradient)

set.seed(1234)
```

# Introduction

Building a mental lexicon is a major achievement in the development of an infant: by storing representations of how familiar words sound and what they mean, an infant is able to make sense of their linguistic input. The foundations of an initial lexicon are in place before their first year of life [@tincoff1999some; @halle1994emergence; @vihman2004cross; @parise2012electrophysiological; @bergelson20126; @bergelson2015early]. This initial lexicon consists of only a few items; mainly words for people, interjections, body parts, and food [@tincoff2012six; @tardif2008baby], but it undergoes rapid growth during the second year of life [@ganger2004reexamining; @goldfield1990early; @bloom2002children; @mcmurray2007defusing]. According to parental reports, the average 15-month-old infant already understands more than 100 words, and by two years, they understand more than 400 words [@frank2021variability]. This accelerated lexical developmental is reflected in infants' trajectories of word recognition: infants recognise familiar words faster and more efficiently as they approach their second birthday [@fernald1998rapid; @fernald2001half; @hurtado2007spoken].

Despite being exposed to a more complex linguistic input, bilinguals show equivalent trajectories of word acquisition and word recognition to their monolingual peers' [@vihman2007onset; @legacy2018vocabulary; @hoff2012dual; @de2014bilingual; @pearson1994patterns; @byers-heinlein2023sometimes]. This is a remarkable deed for two reasons. First, bilingual infants receive a relative impoverished linguistic input in each of their languages compared to monolinguals [@costa2014does]. Second, they face a more complex referential context, since they often learn two labels for each referent, one in each language. The mechanisms that allow bilingual' trajectories of lexical developmental to keep up with monolinguals' are still unclear.

Monolinguals and bilinguals share broadly common word-recognition mechanisms. The first step towards spoken word-recognition is the acoustic-phonetic processing of the speech signal, which results in the activation of lexical representations whose associated phonological forms match the acoustic signal. The activated word-forms then compete for selection: the lexical system uses the bottom-up information provided by the auditory input, together with top-down context---provided by grammatical, supra-segmental, and semantic constraints---to select the candidate lexical representation with the best match. This sequence of events occurs in a *cascaded* fashion. As the recognition system accumulates information about the unfolding acoustic signal, matching phonological representations are activated [@marslen1987functional; @marslen1978processing; @grosjean1980spoken], and in turn activate their associated semantic representations [@neely1977semantic]. By the time lexical selection occurs, many non-selected phonological and semantic representations have been activated, potentially modulating the dynamics of word recognition. 

A considerable amount of evidence for such cascaded account of lexical processing has been provided by studies recording eye movements during word recognition. For example, @allopenna1998tracking designed a word-recognition task in which participants were presented, in each trial, with four objects in the screen. Participants would then listen to a command like "Pick up the beaker; now put it below the diamond", and perform the action. The authors manipulated the phonological relationship between the target object's label (e.g., *casket*), and the label of each of the three distractors. Two of the distractors were phonologically related to the target, sharing onset (e.g., *castle*) or offset (e.g., *basket*), respectively. The other distractor was phonologically unrelated to the target (e.g., *nickel*). After the onset of the auditory target label, phonological distractors attracted participants' eye fixations more than unrelated distractors. Distractors sharing phonological onset did so at an earlier time window than those sharing phonological offset. This time course of distractor fixations suggests that the auditory presentation of the target label activated no only its (subsequently selected) phonological form in participants' lexicon, but also that of phonologically related word-forms, leading to competition for selection between the word forms. These results converge with cascaded models of auditory word recognition, in which the unfolding acoustic input activates multiple phonological word-forms, which in turn compete for selection, like TRACE [@mcclelland1986trace] and Shortlist [@norris1994shortlist].

The developing mental lexicon also seems to operate in this cascaded fashion: like adults, infants activate non-selected lexical representations. For instance, priming paradigms of word recognition have provided a strong body of evidence for the existence of phonological and semantic links between early lexical representations between 18 and 21 months of age.. @arias2009lexical found that infants' spoken word recognition (e.g., *dot*) was interfered by the previous presentation of a semantically related word (e.g., *cat*). Other studies have provided electrophysiological evidence of semantic priming at earlier ages [e.g., @rama2013development; @friedrich2005lexical]. More recent studies have provided evidence for the emergence of such semantic links, even in the absence of visual referents during the experimental task [@willits2013toddlers]. Phonological priming effects have also been reported in the initial lexicon. @mani2010infant and @mani2011phonological created an implicit naming task in which each trial started with the *silent* presentation of a prime picture. Then, a target-distractor picture pair was presented side-by-side. Finally, the target picture's label was presented auditorily. The authors manipulated the phonological overlap between the prime label and the target label, so that in half of the trials both labels were phonologically related, sharing phonological onset (*cat*-*cup*), or phonologically unrelated (*ball*-*comb*). Prime and target-distractor pairs were semantically unrelated. Interestingly, infants showed different target looking preference patterns after phonologically related prime pictures, compared to after phonologically unrelated primes. This suggests that infants implicitly named prime pictures---despite such pictures having been presented in silence---and that the phonology of the resulting labels interacted with subsequent target.

@chow2017spoken provided further evidence of semantic *and* phonological links in the early lexicon. The authors adapted the Visual World Paradigm from @huettig2007tug to explore 24- to 30-month-old toddlers' visual fixation patterns during a word recognition when presented with phonological and semantic distractors. In each trial, the authors presented participants with four semantically and phonologically pictures. Four seconds after pictures onset, a word-form was auditorily presented. The word did not refer to any of the pictures displayed on the screen, but was phonologically related to one of them (both labels shared phonological onset), and semantically related to another one of the pictures (both referents belonged to the same taxonomic category. For instance, participants might be presented with the pictures of a sandwich, a bus, a cat, and a dress. Then they would hear the carrier phrase "Look at the *bee*!". The authors registered participants fixations to the phonological and semantic distractors, and found evidence of a preference for the phonological distractor at earlier stages of the post-naming phase, and a preference for the semantic distractor at later stages of the trial. These results support a cascaded activation account of lexical access during the first stages of lexical development, paralleling previous findings in adults [@allopenna1998tracking].

One of the most interesting pieces of evidence in support of a cascaded account of lexical processing is provided by bilinguals. A critical property the bilingual lexicon is that its representations are accessed in a language-non selective way. Even during monolingual situations, bilinguals access their lexical representations in both of their languages [e.g., @thierry2007brain; @dunabeitia2009masked; @hoshino2008cognate; @strijkers2010tracking; @duyck2007visual]. This is reflected in bilinguals' performance in word recognition and production tasks. For instance, @marian1999activation presented Russian-English bilinguals with a series of trials that started with an auditory instruction like "Poloji marku nije krestika" ["Put the stamp below the cross"]. The instruction would be present in Russian (dominant language for most participants). The target object (*marku* [*stamp*], in this case) would be present in the screen, together with an object whose English label shared phonological onset with the target label (e.g., *mark*), and two unrelated distractors whose labels in Russian and English did not shared phonological relationship with the target label. In line with @allopenna1998tracking's results in monolingual adults, bilinguals fixated the cross-language phonological distractor significantly more than the unrelated distractors after hearing the target label. These findings suggest that bilinguals activated phonologically related word-forms in both Russian *and* English, which ultimately affected their overt visual exploration patterns during the task.

Language non-selectivity in the bilingual lexicon has also been shown to affect top-down lexical access during word production. @costa2000cognate presented Catalan-Spanish bilinguals with a series of pictures of familiar objects. For each object, participants were asked to name it in Catalan (their dominant language). Unbeknownst to participants, the authors manipulated the phonological overlap between the pictures' labels in Catalan and their translations in Spanish. In half of the trials, the labels associated to the pictures were cognates, that is, phonologically similar translation equivalents (e.g., *gat*-*gat* [*cat*]) in Catalan and Spanish, whereas in the other half of the trials the labels were non-cognates (e.g., *taula*-*mesa* [*table*]). Participants named pictures faster in cognate trials than in non-cognate trials. Critically, Spanish monolinguals who completed the same task, naming the pictures in Spanish, showed equivalent naming times in both conditions. These results reveal that bilinguals activated their Spanish phonology, despite performing the naming task exclusively in Catalan: the visual recognition of the presented pictures led to the activation of its associated phonological forms in both languages in parallel, which influenced the subsequent the dynamics of word production. This differential processing of cognates and non-cognates has generated a large body of literature providing support for the language-non selective account of lexical access [e.g., @degrootandnas1992; @schwartzkroll2007; @dijkstragrainger1999; @dufour1995matching; @schwartz2007reading; @dijkstra2010cross].

Further evidence in favour of the parallel retrieval of phonological forms during bilingual word production is given by implicit naming paradigms. Using an adaptation of @mani2010infant's task, @von2014bilinguals found that bilingual adults generate implicit labels in both labels for visually fixated pictures. The authors presented German-English bilinguals with 120 prime-target pairs. Primes were presented as familiar pictures in silence. After prime picture offset, target words were presented auditorily. Participants' N400 ERP components were recorded from target word onset. The authors manipulated the phonological relationship between the prime and target word-forms within participants' L1 and across L1 and L2. When prime and target were phonologically identical (*Affe*-*Affe*, German for *monkey*), or similar (*Fahne*-*Sahne*, German for *flag* and *ice-cream*, respectively) within L1, participants' showed a reduced N400 amplitude, compared to when prime and target were phonologically unrelated (*Messer*-*Seil*, German for *knife* and *rope*). Critically, a similar effect was found when prime and target were phonologically related through translation (*Rustsche*-*Kleid*, German for *slide* and *dress*). This suggests that participants activated prime labels in both languages in parallel, and that both labels impacted the dynamics of target words recognition.

Phonological priming paradigms have also revealed language non-selectivity in the initial bilingual lexicon [@poarch2012cross; @bosma2020cognate; @singh2014one; @jardak2019labels; @von2019impact]. For instance, @von2012language found evidence of cross-language phonological priming in 21- to 42-months-old children learning German and English. At the beginning of each trial, the authors auditorily presented an English prime word embedded in a carrier phrase. Then, the target label was also auditorily presented in German. Finally, the target and distractor pictures were presented side-by-side. In some trials, the auditorily presented prime and target labels were phonologically related (e.g., *slide*-*Klide* [*dress*]). In some other trials, both labels were phonologically related through translation: the auditorily presented prime label did *not* overlap with the target label (e.g., *leg*-*Stein* [*stone*]), but its translation in German did [*Bein*]. In the rest of the trials prime and target labels were phonologically unrelated in both languages. The authors found a facilitation effect of cross-language priming when both prime and target auditory labels overlapped phonologically, as revealed by a stronger target picture looking preference, as compared to that of unrelated trials. Interestingly, participants showed a weaker target preference in priming through translation trials.

Using a similar task, @floccia2020translation provided conflicting results to those of @von2012language. The authors tested 27-month-old simultaneous bilinguals in a cross-language priming paradigm. The two languages participants were learning were English, and an Additional Language (either Cantonese, Dutch, French, German, Greek, Italian, Mandarin, Polish, Portuguese, or Spanish). A prime word was embedded at the end of a carrier sentence that participants listened to. Then, participants were presented with the auditory label of the target word, and two pictures were shown side-by-side, the target picture, and a distractor picture. In Study 1, the target word was the translation equivalent of the prime word (*cheese*-*fromage* for an English-French bilingual). In other trials, prime and target words were unrelated (e.g., *sock*-*fromage*). There was no phonological overlap at onset between prime and target or distractor labels in either of the two conditions. Participants were presented with two blocks of related and unrelated trials. In one of the blocks, participants were presented with English prime words and Home Language target words (*sock*-*fromage*), and in the other block they were presented with Additional Language prime words and English target words (*fromage*-*sock*). The authors found a cross-linguistic priming effect, in which participants showed a stronger preference for the target picture in the post-naming phase in related trials, as compared to unrelated trials. This effect was found regardless of the language of the prime (English or Home Language), or the dominance of the language of the prime (i.e., whether the prime word belonged to the dominant language). Target preference exceeded chance level in both conditions, pointing to successful word recognition regardless of the experimental manipulation.

In a second study, @floccia2020translation, the authors tested an equivalent group of bilinguals in a cross-linguistic semantic priming task, in which participants were primed with words in English, and tested with semantically related words in the Additional Language (or vice versa). This time, prime and target words were not translation equivalents, but rather words whose referents shared semantic features. Results paralleled those from their first study. The authors found a strong semantic priming effect across languages, regardless of the language of the prime or the dominance of the language of the prime. Overall, these results suggest that bilinguals recognised target words in both conditions, and benefited from a cross-linguistic facilitation effect from translation equivalents. The symmetry of such effect across both languages suggests that participants lexical access occurred in parallel for the two members of translation equivalents. As the authors noted, the stimuli set of their second study contained multiple cognates due to design constraints, especially in languages with a larger degree of phonological overlap across translation pairs (e.g., English and Dutch). The authors run *post hoc* analyses exploring the effect of cognateness on the cross-language priming effects found, which showed no evidence of a cognateness effect on target preference. However, the cross-language effects reported in this second study might still reflect the confounding influence of cognateness, as the resulting effect might have been generated by an interaction by the phonological and semantic overlap between cognate primes and targets.

Previous studies had reported similar cross-language priming effects. The conclusion that such effects occurred at the lexical level, as opposed to being the result of mere acoustic similarity, was nonetheless unwarranted. In studies where the auditorily presented prime label was phonologically overlapped with the target label [e.g., @marian1999activation], priming effects might be explained by the activation of the phonological segments of the prime label, which facilitated the recognition of the target label, which shared such phonological segments. The priming-through-translation results by @von2012language and @floccia2020translation demonstrate that cross-language activation in the developing lexicon occurs at the lexical level. Since participants' target recognition was modulated by the non-presented prime translation, such an inhibitory effect must stem at the lexical level. The fact that phonological priming facilitated the recognition of the target words (both labels belong to different languages), and that phonological priming through translation interfered target recognition (translation and target belong to the same language) points to cross-language connections playing an excitatory role, and to within-language connections playing an inhibitory role.

The priming-through-translation paradigm provides a convenient framework to investigate parallel activation in the bilingual lexicon, but its design prevents strong interpretations about the extent to which parallel activation is the result of the experimental manipulations of interest. The presentation of primes and targets in both languages in the same task---or even within the same trial---presents participants with a bilingual context that might lead to a higher baseline activation of lexical representations of both languages, contrary to a context in which parallel activation results from genuine cross-language activation [@grosjean1997bilingual]. Another limitation of priming-through-translation studies is that, when prime words are auditorily presented, the task constrains the naming context to one of the languages. The consequence is a potential asymmetry between the activation dynamics of the lexical representation of both labels: while the participant has heard one label, the activation of its translation must spread across semantic links. This might lead to a lower strength, and higher latency of the activation of the translation. 

In the present study, we put the hypothesis that bilingual infants activate both languages in parallel to a more severe test. We used an extension of the implicit naming paradigm by @mani2010infant, in which participants were primed with a picture presented in silence. This provides participants with an exclusively monolingual context in which they perform the task in only one of their languages. If differences in task performance are found associated with covert manipulation of cross-linguistic properties of the stimuli, such differences may be more conclusively linked to the parallel activation of lexical representations in participants' lexicons. By using an implicit naming paradigm, the activation of the lexical representation of the prime may occur in both languages as the result of lexical retrieval after visual recognition of the prime picture, since participants are not presented with any auditory prime label. This contrasts with priming through-translation tasks, in which the prime label in one of the languages is directly activated by auditory word recognition, while the label of its translation is expected to be activated as the result of cascaded activation spreading through their semantic association. This may lead to an asymmetry in the time course of the activation of both labels, which might in turn affect the dynamics of phonological priming in subsequent word recognition.

In summary, the present study was aimed at investigating the developmental trajectories of word recognition in bilingualism, and how (if at all) language non-selectivity shapes them. Using an adaptation of [@mani2010infant], we presented 20- to 32-months-old toddlers with a cross-linguistic phonological priming paradigm. We manipulated both the phonological overlap between the prime word and target word, and the cognate status of the prime word, to test the hypothesis that bilingual toddlers implicitly name familiar pictures in both languages. If so, phonological priming effect should be larger when the prime label in both languages shares phonological overlap with the target words (*Related/Cognate* condition), compared to when only the label of the prime in one language overlaps phonologically with the target (*Related/Non-cognate*). We also also included a control *Unrelated* condition in which regardless of the cognate status of the prime word, neither of the prime labels in any language overlapped phonologically with the target word.

To test cross-language priming effects, we registered participants' time course of target picture looking in the three conditions. We predicted within-language priming effects, revealed as differences in target looking between *Unrelated* and *Related/Non-cognate* trials, as previously reported [@von2012language, @mani2010infant, and @mani2011phonological]. We further predicted cross-language priming effects, revealed as differences in target looking between *Related/Non-cognate* and *Related/Cognate* trials. 
This effect should be stronger in bilinguals, if not completely absent in monolinguals, who should be indifferent to the phonological properties of the stimuli in the other language.

Following previous studies associating the emergence of phonological links between lexical items to the size of children's vocabulary [@mani2011phonological], we also compared the explanatory power of two indices of receptive vocabulary size (dominant-language vocabulary size and total vocabulary size) against participants' age. We used parental reports of vocabulary checklist to estimate participants' vocabulary size. We computed vocabulary size in the dominant language (L1 vocabulary) as the proportion of words in the dominant language reported by their caregivers to be understood by the child in a vocabulary checklist, and the total vocabulary size as the proportion of words reported as understood by the child in *both* languages. If within-language or cross-language phonological associations in the lexicon emerge as a function of vocabulary growth, the size of the differences between conditions should larger in participants with larger vocabulary sizes. A model including participants' L1 or total vocabulary size as a predictor should then show better performance than a model that instead includes participants age. In the case of bilinguals, if the emergence of cross-language priming effects is associated with the growth of their vocabulary is *both* languages, a model including total vocabulary size as a predictor should perform better than a model that includes L1 vocabulary size. The opposite pattern would indicate that cross-language priming effects emerge as a function of the growth of L1 vocabulary.

Testing bilingual populations at early ages presents the limitation that the vocabulary size of participants' non-dominant language is relatively small, which substantially constrains the number of items that can be used to as stimuli in the task, as well as the number of trials that can be constructed to generate stimuli lists. In order to circumvent the problem of limited vocabulary knowledge in the non-dominant language, tested participants only in their dominant language, therefore directing our research to the question of how possessing a second language affects processing of the dominant one in bilingual toddlers [@costa2014does provide a fuller description of the advantages of such approach].

# Study 1

## Methods

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/ckydb/)) and GitHub ([https://github.com/gongcastro/cognate-priming](https://github.com/gongcastro/cognate-priming)) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I. Before every testing session, caregivers were asked to read and sign an informed consent form, and were given a token of appreciation at the end of it.

### Participants

```{r}
#| label: participants-numbers
participants_bcn <- filter(participants, location=="Barcelona")

n_participants_total <- nrow(distinct(participants_bcn, child_id))
n_sessions_total <- count(participants_bcn)

n_participants_lp <- participants_bcn |> 
	distinct(lp, child_id) |> 
	count(lp)

n_participants_lp_sex <- participants_bcn |> 
	distinct(lp, sex, child_id) |> 
	count(lp, sex)

n_participants_sessions <- participants_bcn |> 
	count(child_id, name = "n_sessions") |> 
	count(n_sessions) |> 
	group_split(n_sessions) |> 
	set_names(paste0("session_", 1:3)) |> 
	map("n")

n_sessions_lp <- participants_bcn |> 
	summarise(across(age, lst(mean, sd, min, max)), n = n(), .by = c(lp)) |> 
	mutate(across(age_mean:age_max, \(x) round(x, 2))) |> 
	group_split(lp) |> 
	set_names(janitor::make_clean_names(unique(participants_bcn$lp)))

n_sessions_dominance_lp <- count(participants_bcn, lp, test_language) |> 
	group_split(test_language) |> 
	set_names(c("catalan", "spanish")) |> 
	map(group_split, lp) 

n_sessions_lp <- participants_bcn |> 
	count(lp) |> 
	group_split(lp) |> 
	set_names(c("monolingual", "bilingual")) |> 
	map("n")
```


We collected data from `r n_participants_total` children living in the Metropolitan Area of Barcelona (Spain), tested at the Laboratori de Recerca en Infància at the Universitat Pompeu Fabra. Families were recruited from maternity rooms in private hospitals and social media, and contacted via phone when the child's age spanned between 20 and 32 months. From the `r n_participants_total` children that participated, `r n_participants_sessions$session_1` participated once, `r n_participants_sessions$session_2` participated twice, and `r n_participants_sessions$session_3` participated three times. Recurrent participants were tested with at least XXX months of difference. We gathered a total of `r n_sessions_total$n` testing sessions. Participants were divided into monolinguals and bilinguals based on their relative degree of exposure to Catalan and Spanish, estimated using the Language Exposure Questionnaire [LEQ, @bosch2001evidence]. We categorised participants as monolingual if exposed to more than 80% or more of the time to their dominant language, and as bilingual otherwise. `r english::Words(n_participants_lp$n[1])` of the participants were categorised as monolinguals (`r n_participants_lp_sex$n[1]` female, `r n_participants_lp_sex$n[2]` male) and `r n_participants_lp$n[2]` as Catalan/Spanish bilinguals (`r n_participants_lp_sex$n[3]` female, `r n_participants_lp_sex$n[4]` male) (see @tbl-participants for a detailed summary of participants' age and language profile). Participants' vision was normal, none used glasses or any other type of vision corrector.


```{r tbl-participants}
#| label: tbl-participants
#| tbl-cap: "Demographic and linguistic profile of testing sessions."
participants_bcn |> 
	unnest_wider(doe) |> 
	select(-matches("others")) |> 
	arrange(lp, test_language) |> 
	summarise(n = n(),
			  n_participants = n_distinct(child_id),
			  across(age, lst(mean, sd, min, max)),
			  across(matches("doe_"), lst(mean, sd)),
			  # age_vctr = list(age),
			  .by = c(lp, test_language)) |> 
	gt(rowname_col = "test_language",
	   groupname_col = "lp") |> 
	# gt_plt_dist(age_vctr, type = "histogram") |>
	fmt_number(is.numeric, decimals = 1) |> 
	fmt_integer(c(n, n_participants)) |> 
	fmt_number(matches("doe"), decimals = 1, scale_by = 100) |> 
	cols_merge(c(age_min, age_max), 
			   pattern = "{1} ({2})") |> 
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(doe_spanish_mean, doe_spanish_sd), 
			   pattern = "{1} ({2})") |> 
	tab_spanner("Age (months)", matches("age")) |> 
	tab_spanner("Catalan", matches("catalan")) |> 
	tab_spanner("English", matches("english")) |> 
	tab_spanner("Spanish", matches("spanish")) |> 
	tab_spanner("Degree of Exposure (%)", matches("doe")) |> 
	cols_label(n = "N sessions",
			   n_participants = "N participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   # age_vctr = "Distribution",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())

col_names <- c("Group",
			   "Language",
			   "Sessions", 
			   "Participants",
			   "M (SD)",
			   "Range",
			   "M (SD)",
			   "M (SD)",
			   "M (SD)")
participants_bcn |> 
	unnest_wider(doe) |> 
	select(-matches("others")) |> 
	arrange(lp, test_language) |> 
	summarise(n = n(),
			  n_participants = n_distinct(child_id),
			  across(age, lst(mean, sd, min, max)),
			  across(matches("doe_"), lst(mean, sd)),
			  # age_vctr = list(age),
			  .by = c(lp, test_language)) |> 
	mutate(age_mean = sprintf("%.1f (%.1f)", age_mean, age_sd),
		   age_range = sprintf("%.1f-%.1f", age_min, age_max),
		   doe_catalan = sprintf("%.1f (%.1f)", doe_catalan_mean*100, doe_catalan_sd*100),
		   doe_spanish = sprintf("%.1f (%.1f)", doe_spanish_mean*100, doe_spanish_sd*100),
		   doe_english = sprintf("%.1f (%.1f)", doe_english_mean*100, doe_english_sd*100)) |> 
	select(lp:age_mean, age_range, doe_catalan, doe_spanish, doe_english) |> 
	kable(digits = 1, 
		  format = "latex",
		  booktabs = TRUE,
		  col.names = col_names,
		  align = "llrrrrrrrrrrrr"
	) |> 
add_header_above(c(
	" " = 4,
	"Age (months)" = 2,
	"Degree of exposure (%)" = 8
)) |> 
	row_spec(nrow(tbl_data) - 1,
			 hline_after = TRUE
	) 
	add_footnote("For participants exposed to Catalan and Spanish exclusively, the proportion of Catalan exposure shown in the table is complementary to the degree of exposue to Spanish. For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10\\%. Participants with 100\\%, 0\\%, and 50\\% Catalan exposure would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.", notation = "none", escape = FALSE)
```

### Vocabulary size

```{r vocab-values}
vocabulary_bcn <- vocabulary |> 
	inner_join(select(participants_bcn, session_id, lp, age_group),
			   by = join_by(session_id))

vocab_means <- vocabulary_bcn |> 
	summarise(across(c(total_prop, l1_prop), lst(mean, sd)),
			  n = n(),
			  .by = c(age_group, lp)) %>%
	mutate(across(matches("prop"), scales::percent)) %>%
	split(paste0("age_", gsub(" months", "", .$age_group))) |> 
	map(\(x) split(x, tolower(x$lp)))

n_imputed <- table(vocabulary_bcn$is_imputed)[2]
prop_imputed <- scales::percent(n_imputed/nrow(vocabulary_bcn))
n_pool <- nrow(bvq_data$vocabulary)
```

We collected vocabulary data using parental responses to the Barcelona Vocabulary Inventory [BVQ, @garcia-castro2023bvq], an online vocabulary checklist inspired in several adaptations of the the Communicative Developmental Inventory [CDI, @fenson1994variability] developed to assess the vocabulary size of Catalan-Spanish bilingual toddlers. Families received a link to the BVQ immediately after each experimental session, and were given two weeks to fill it. We calculated two measures of receptive vocabulary size for each participant: *L1 vocabulary size* (proportion of words  reported as acquired in the checklist of the dominant language), and *total vocabulary size* (proportion of the words in both checklists reported as acquired). `r n_imputed` (`r prop_imputed`) Families failed to provide a complete response to the BVQ within the two-week time limit, or did not provide a successful response to the questionnaire. For missing questionnaire responses, we imputed the vocabulary size of the participant using single imputation, using the vocabulary size scores of a pool of `r n_pool` additional participants for which a successful response for the questionnaire had been gathered. We used participants age in months and their language profile (monolingual/bilingual) as predictors. We used the `mice` R package [@van2011mice] to perform imputation using the Bayesian linear regression method. 

Overall, there was a substantial increase in both total and L1 vocabulary sizes associated to age (see @fig-vocabulary). Monolinguals showed overall higher L1 vocabulary sizes than monolinguals: they were reported to understand an average of `r vocab_means$age_21$monolingual$l1_prop_mean` (*SD* = `r vocab_means$age_21$monolingual$l1_prop_sd`) of the words at around 21 months, `r vocab_means$age_25$monolingual$l1_prop_mean` (*SD* = `r vocab_means$age_25$monolingual$l1_prop_sd`) at 25 months, and `r vocab_means$age_30$monolingual$l1_prop_mean` (*SD* = `r vocab_means$age_30$monolingual$l1_prop_sd`) at 30 months, while bilinguals  were reported to understand an average of `r vocab_means$age_21$bilingual$l1_prop_mean` (*SD* = `r vocab_means$age_21$bilingual$l1_prop_sd`) of the words at around 21 months, `r vocab_means$age_25$bilingual$l1_prop_mean` (*SD* = `r vocab_means$age_25$bilingual$l1_prop_sd`) at 25 months, and `r vocab_means$age_30$bilingual$l1_prop_mean` (*SD* = `r vocab_means$age_30$bilingual$l1_prop_sd`) at 30 months. Total vocabulary sizes showed the opposite pattern, with monolinguals infants being reported to understand `r vocab_means$age_21$monolingual$total_prop_mean` (*SD* = `r vocab_means$age_21$monolingual$total_prop_sd`) of the words at around 21 months, `r vocab_means$age_25$monolingual$total_prop_mean` (*SD* = `r vocab_means$age_25$monolingual$total_prop_sd`) at 25 months, and `r vocab_means$age_30$monolingual$total_prop_mean` (*SD* = `r vocab_means$age_30$monolingual$total_prop_sd`) at 30 months, and bilinguals being reported to understand `r vocab_means$age_21$bilingual$total_prop_mean` (*SD* = `r vocab_means$age_21$bilingual$total_prop_sd`) of the words at around 21 months, `r vocab_means$age_25$bilingual$total_prop_mean` (*SD* = `r vocab_means$age_25$bilingual$total_prop_sd`) at 25 months, and `r vocab_means$age_30$bilingual$total_prop_mean` (*SD* = `r vocab_means$age_30$bilingual$total_prop_sd`) at 30 months.

```{r fig-vocabulary}
#| label: fig-vocabulary
#| fig-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Grey lines connect vocabulary size scores from the recurrent participants. For visualisation purposes, participants were group into categorical age groups at 21, 25, 30 months, the ages around which most participants were tested. Only participants included in the final data set are shown."
#| fig-width: 6
#| fig-height: 3.5
vocabulary_bcn |> 
	inner_join(filter(attrition_participants, is_valid_participant)) |> 
	pivot_longer(c(total_prop, l1_prop),
				 names_to = "measure",
				 values_to = "prop",
				 names_transform = \(x) gsub("_prop", "", x)) |> 
	mutate(measure = if_else(measure=="l1", 
							 "L1 vocabulary", 
							 "Total vocabulary"),
		   is_imputed = if_else(is_imputed, "Imputed", "Observed") |> 
		   	factor(levels = c("Observed", "Imputed"))) |> 
	ggplot(aes(age_group, prop,
			   colour = lp,
			   fill = lp)) +
	facet_wrap(~measure) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1/4, linewidth = 3/4) +
	geom_dots(aes(side = if_else(lp=="Monolingual", "left", "right")),
			  dotsize = 2,
			  scale = 0.5,
			  alpha = 3/4) +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 linewidth = 3/4,
				 width = 0.15,
				 colour = "black",
				 show.legend = FALSE,
				 position = position_dodge(width = 1.15)) +
	stat_summary(fun = mean,
				 geom = "point",
				 size = 2,
				 colour = "black",
				 show.legend = FALSE,
				 position = position_dodge(width = 1.15)) +
	# geom_point(alpha = 1) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(labels = scales::percent,
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```



### Stimuli

```{r stimuli-values}
n_words <- length(unique(with(stimuli, unlist(prime, target, distractor))))
```

We used `r n_words` distinct words included in the BVQ to create the stimuli lists. We created six stimuli lists: three in Catalan, and three in Spanish. Each list contained 32 trials, each involving a prime-target-distractor group. Each word played a role as either prime, *or* as target and distractor across the three lists in their corresponding language. For instance, the Catalan word *cadira* appeared as *prime* in the three lists, but never as *target* or *distractor*; the Catalan word *bici* appeared as *target* and *distractor* across the three lists, but never as a prime. Target-distractor pairings were held constant across the three lists in each language. For instance, in all Catalan lists the word *bici* was paired with the word *porta*. Target-distractor pairings were also yoked, so that each member of the same target-distractor pair appeared once as target and once as distractor in each list. For instance, the *bici*-*porta* paired appeared twice in each of the three Catalan lists: once with *bici* as target and *porta* as distractor, and once with *porta* as target and *bici* as distractor. This counterbalancing avoided participants encountering looking at the target word guided solely by that word having being named in a previous trial. Finally, prime words appeared only once in each list: each target-distractor pair was associated with a different prime word in both appearances. In each list, the same prime word was presented alongside a different target-distractor pair. For instance, the Catalan prime word *barret* was presented with the *bici*-*porta* target-distractor pair in one list, with the *bici*-*porta* pair in another list, and with *berenar*-*amanida* in the remaining list. To avoid competition between target and distractor pictures, semantically related target-distractor pairs were perceptually distinct [@floccia2020translation; @arias2009lexical]. The order of the trials was randomised across experimental session, so that each time a participant was tested, the order in which the prime-target-distractor was presented was randomised. Each participant was randomly assigned to one of the three lists in the corresponding (dominant) language, and always the same list across their experimental sessions in the case of a recurrent participant.

In 16 of the 32 trials of the same list (henceforth *related* trials), the prime and the target words were phonologically related, sharing phonological onset (at least first phoneme). In the other 16 trials (*unrelated* trials), prime and target did not share phonological onset. 8 of the 16 *related* trials included a cognate prime (*cognate* trials), and the other 8 included a non-cognate trials (*non-cognate* trials). A prime word was considered cognate if its Catalan and Spanish translation shared phonological onset. Especial attention was paid to avoiding semantic or taxonomic relationships between prime and target words, and between prime and distractor words. Target and distractor word pairs were phonologically unrelated (did not share phonological onset). Some of them shared semantic features or a taxonomic relationship. This is the case of words associated with especially salient referents such as animals or food. To avoid infants guiding their gaze to these objects based on their saliency, we paired animals and food items together. The position of the target and distractor pictures (right or left) for each target-distractor pair was alternated, so that in one list the target would appear on the left, in another list it would appear on the right, and so on.

We examined the overall equivalence of the three trial types by comparing them across three variables relating to the target word: lexical frequency, word prevalence, animacy. @tbl-stimuli shows a detailed summary of the stimuli properties, broken down by trial type and testing language. Lexical frequencies were extracted from the Catalan and Spanish corpora of the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex]. We defined word prevalence as the proportion of same-aged infants who were reported to understand the word in the BVQ database.


#### Auditory stimuli

The auditory stimuli were natural exemplars of the selected target words, spoken by Central Catalan-Spanish proficient bilingual female speaker who was instructed to pronounce each word in a toddler-directed manner. Recordings were made with an Audio-Tecnica 328 microphone (AT2050) at a sampling rate of 44100 Hz, in a soundproof room at the *Laboratori de Recerca en Infancia* at University Pompeu Fabra. We used the Audacity and Praat [@boersma2001speak] software packages to record and edit the audio files. The speaker was presented with a list of words in Catalan. The order of the words was pseudo-randomised, and each word was produced three times in a row before moving to the next word in the list. After going through all the words in the list, the speaker went through the word list again generating three tokens for each word, now in an inverse order (from bottom of the list to the top). We then repeated the same procedure for the list of Spanish words. The resulting audios were manually chunked into individual word-forms. For each of the six tokens produced for each word, the most adequate was selected for further processing. The audios were then transformed to stereo by duplicating them into two channels, denoised, and finally normalised. The mean duration of the final audios was `r round(mean(stimuli$duration[stimuli$test_language=="Catalan"]), 2)` (*SD* = `r round(sd(stimuli$duration[stimuli$test_language=="Catalan"]), 2)`) and `r round(mean(stimuli$duration[stimuli$test_language=="Spanish"]), 2)` (*SD* = `r round(sd(stimuli$duration[stimuli$test_language=="Spanish"]), 2)`) seconds for the Catalan and Spanish lists.

To make the pronunciation of the words as familiar as possible to each infant, we generated additional pronunciation variants for some words in Catalan and Spanish. Catalan words involving the /\textipa{L}/ phoneme in their Central Catalan variant (e.g., /\textipa{'Lu.n@}) were also recorded with such phoneme replaced by /j/ (e.g., /\textipa{'ju.n@}), a phonological process common in the Metropolitan Area of Barcelona. Spanish words involving the /\textipa{T}/ phoneme were also generated replacing such phoneme with /\textipa{s}/ to better accommodate Latin variants of Spanish. Before every experimental session, caregivers were asked to utter three written words involving the /\textipa{L}/ phoneme (in the case of participants tested in Catalan) or the /\textipa{T}/ phoneme (in the case of participants tested in Spanish). Each token contained the critical phoneme at onset, inter-vocalic position, and coda. The experimenter assigned the participant to the Catalan or Spanish stimuli list involving the closest variant to that of caregivers'.


#### Visual stimuli

Each word was depicted by a realistic phonographic representation of a typical exemplar. Image backgrounds were removed from the original pictures using the GNU Image Manipulation Program (GIMP), resized to a rectangle of a maximum of 400 pixels hight or wide, and finally placed in the centre of a 50% grey rectangle square of 500$\times$500 pixels. The final stimuli had a resolution of 72 dpi. When presented in the eye-tracker screen, the areas of interest (AOI) occupied an area of 13.23$\times$13.23 cm (11.613º visual angle from participants' perspective).

```{r tbl-stimuli}
#| label: tbl-stimuli
#| tbl-cap: "Summary of stimuli properties by trial type."
stimuli |> 
	summarise(across(c(matches("familiarity_|freq_|animate_"), duration),
					 tibble::lst(mean, sd, min, max)),
			  .by = c(test_language, trial_type)) |> 
	select(-matches("familiarity_se"), -matches("prime"),
		   -c(is_animate_target_sd,
		      is_animate_target_min,
		      is_animate_target_max),
		   -matches("min"), -matches("max")) |> 
	gt(groupname_col = "test_language",
	   rowname_col = "list") |> 
	fmt_number(is.numeric,
			   decimals = 1) |> 
	fmt_number(matches("familiarity|animate"),
			   scale_by = 100,
			   decimals = 1) |> 
	cols_merge(c(freq_target_mean, freq_target_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(familiarity_target_mean, familiarity_target_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(duration_mean, duration_sd),
			   pattern = "{1} ({2})") |> 
	cols_label(trial_type = "",
			   familiarity_target_mean = "Prevalence (%)",
			   freq_target_mean = "Frequency",
			   is_animate_target_mean = "Animacy (%)",
			   duration_mean = "Duration (s)") 
```


### Procedure

Testing took place in a sound-proof room. Participants sat on their caregivers' lap in a dimly lit testing booth while the experimenter conducted the experiment from outside. Caregivers were instructed to keep their eyes shut (to avoid recording their gaze, instead of the participant's), to be still, and to avoid interacting with the participant verbally or non-verbally. Participants sat at approximately 65 cm from the eye-tracker and a XX-in screen of $1929\times1080$ screen resolution. We used a custom Matlab XXXX script using the PsychToolbox-3 extension [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox] to present the stimuli, and the Tobii Analytics SDK 3.0 to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 5-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented. We set a 55% grey background for the calibration and stimuli presentation. Auditory stimuli were presented through two loudspeakers located behind the screen, one to each side. The experimenter monitored the experimental from outside the room using a centrally located video camera place above the screen. After a successful calibration the experimenter triggered the onset of the first trial. Trials were presented uninterruptedly and  without intervention of the experimenter until the 32 trials were presented, or the experimental session had to be stopped because of the participant's behaviour.

![Experimental task design with examples in Catalan. In each trial, the prime image is presented in silence for 1,500 ms. The the auditory target label is presented, and finally the target and distractor pictures are presented side-by-side for 2,000 ms. In cognate trials (*n* = 8), Catalan *and* Spanish prime labels shared phonological onset with the target label. In non-cognate trials (*n* = 8), only the Catalan prime label shared phonological onset with the target label. In unrelated trials (*n* = 32), none of the prime labels shared phonological onset with the target label.](_assets/img/design.png){fig-align="center"}

Each trial started with the presentation of an attention getter for 3,000 milliseconds. Then, the prime picture was presented in silence in the centre of the screen for 1,500 milliseconds. Fifty milliseconds after the offset of the prime image, an auditory label was played from the loudspeakers and, 700 milliseconds after the onset of the auditory label, the target and distractor pictures were presented side-by-side during 1,000 milliseconds until the end of the trial. After this, the attention getter of the next trial was immediately presented. Each experimental session took approximately 10 minutes.


### Data analysis

#### Data processing

```{r dataset}
attrition_bcn <- filter(attrition, location=="Barcelona")
n_trials <- nrow(attrition_bcn)
n_sessions <- n_distinct(attrition_bcn$session_id)
n_participants <- n_distinct(attrition_bcn$child_id)
n_trials_valid <- nrow(filter(attrition_bcn,
							  is_valid_participant,
							  is_valid_trial))
n_sessions_valid <- sum(distinct(attrition_bcn,
								 session_id,
								 .keep_all = TRUE)$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition_bcn[attrition_bcn$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition_bcn$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_bcn$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_bcn$is_valid_gaze_test_each)
n_exc_participants <- sum(!distinct(attrition_bcn,
									session_id,
									.keep_all = TRUE)$is_valid_participant) 
n_exc_participants_vocab <- sum(!distinct(attrition_bcn,
										  session_id,
										  .keep_all = TRUE)$is_valid_vocab_size) 

n_longitudinal <- attrition_bcn |>
	filter(is_valid_participant,
		   is_valid_trial) |>
	distinct(child_id, session_n) |> 
	count(child_id, name = "times") |> 
	count(times)
```

We defined a time windows of interest from 200 ms after target and distractor pictures onset until the end of the trial at 2,000 ms when  both pictures disappeared from screen. To avoid modelling fixations driven by processes other than auditory word recognition [@fernald1998rapid; @fernald2001half]. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach], with a maximum of 20 maximum consecutive missing samples being interpolated (an equivalent of 166.67 ms).

We gathered data from `r format(n_trials, big.mark = ",")` trials from `r n_sessions` testing sessions, generated from `r n_participants` distinct participants. We excluded trials in which participants failed to provide 50% valid eye-tracking samples (equivalent to 750 ms) during the prime phase (*n* = `r format(n_exc_prime, big.mark = ",")`) or 50% valid samples (equivalent to 1,000 ms) during the target-distractor phase (*n* = `r format(n_exc_test, big.mark = ",")`). We also excluded trials in which participants did not provide at least 5% of valid samples (equivalent to 100 ms) of target or distractor looking in the test phase (*n* = `r format(n_exc_test_each, big.mark = ",")`) [see @floccia2020translation for a similar approach].

After trials that matched any of the aforementioned exclusion criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the  *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 10% (*n* = `r n_exc_participants_vocab`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. @tbl-attrition-trials shows a detailed description of the trial attrition.


```{r tbl-attrition}
#| label: tbl-attrition
#| tbl-cap: "Participant-level and trial-level sample size after applying inclusion criteria. The number of excluded participants and trials is indicated between parentheses."
expanded <- expand(attrition_bcn, lp, trial_type, is_valid_trial)

n_lp <- attrition_bcn |> 
	distinct(is_valid_participant, lp, session_id) |> 
	count(is_valid_participant, lp) |> 
	pivot_wider(names_from = is_valid_participant,
				values_from = n,
				names_repair = janitor::make_clean_names) |> 
	mutate(n_participants_true = true,
		   n_participants_false = false) |> 
	select(lp, true, false)

attrition_bcn |> 
	summarise(n_trials_true = sum(is_valid_trial),
			  n_trials_false = sum(!is_valid_trial),
			  .by = c(lp, trial_type)) |> 
	arrange(desc(lp)) |> 
	pivot_wider(id_cols = lp,
				names_from = c(trial_type),
				values_from = matches("trials"),
				names_repair = janitor::make_clean_names) |> 
	left_join(n_lp, by = join_by(lp)) |> 
	select(lp, true, false,
		   matches("n_trials")) |> 
	gt(rowname_col = "lp") |> 
	tab_spanner("Participants", matches("participant")) |> 
	tab_spanner("Trials", matches("trial")) |> 
	fmt_integer(everything()) |> 
	cols_merge(c(true, false),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_cognate,
				 n_trials_false_cognate),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_non_cognate,
				 n_trials_false_non_cognate),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_unrelated, 
				 n_trials_false_unrelated),
				 pattern = "{1} ({2})") |> 
	cols_label(true = "Participants", 
			   n_trials_true_cognate = "Cognate",
			   n_trials_true_non_cognate = "Non-cognate",
			   n_trials_true_unrelated = "Unrelated") |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	grand_summary_rows(fns = list(label = md("*N*")) ~ sum(.),
					   fmt = ~ fmt_integer(.))
```

#### Modelling approach

```{r data-analysis}

r_info <- sessionInfo()

r_version <- glue::glue(r_info$R.version$major,
						r_info$R.version$minor,
						.sep = "."
)
stan_version <- "2.33.0"

rope_interval <- c(-0.1, +0.1)
```

We used Bayesian Hierarchical General Additive Models (HGAMs) to model our data and test our hypotheses [@pedersen2019hierarchical]. First, we defined our response variable as the empirical logit of target looking using @eq-elogit across all eye-tracking samples for each each participant in each condition [@barr2008analyzing; @agresti2012categorical]. We used a Gaussian distribution to model our response variable. To test our hypotheses, we then included *Condition*, *Group*, and either *Age* ($\mathcal{M}_0$), *L1 vocabulary* ($\mathcal{M}_1$), or *Total vocabulary* ($\mathcal{M}_2$) as fixed effects in the model. We also added all two-way and three-way interactions between the predictors. We set two *a priori* contrasts for the *Condition* predictor: one comparing *Unrelated* and *Related/Non-cognate* trials (sum-coded as `-0.5` and `+0.5`, with *Related/Cognate* trials coded as `0`), and another comparing *Related/Non-cognate* and *Related/Cognate* trials (sum-coded as `-0.5` and `+0.5`, with *Unrelated* trials coded as `0`). We also set two *a priori* contrasts for the *Group* predictor: one comparing *Monolingual (ENG)* with *Monolingual (CAT/SPA)* participants (sum-coded as `-0.5` and `+0.5`, with *Bilingual (CAT-SPA)* participants coded as `0`), and another comparing *Monolingual (ENG)* with *Bilingual (CAT-SPA)* participants (sum-coded as `-0.5` and `+0.5`, with *Monolingual (CAT/SPA)* participants coded as `0`). Before entering the model, the numeric predictors *Age*, *L1 vocabulary*, and *Total vocabulary* were standardised by subtracting from each observation the mean of the predictor, and dividing the result from the standard deviation of the predictor.

$$
\eta' = \ln\bigg(\frac{\text{Target} + 0.5}{ \text{Distractor} + 0.5}\bigg)
$$ {#eq-elogit}

We included *Session*---which indexes individual testing sessions that may belong to the same participant---as grouping variable, nested within *Participant* grouping variable---which indexes a distinct participant. This nested random effects structure incorporates the longitudinal design of data collection, in which multiple participants were tested more than once at different ages. We added by by-session intercepts and *Condition* slopes, and by-participant intercepts and *Age* slopes. To model the time course of target looking across time bins, we included B-splines for the main effect of *Time*, and one for an adjustment of the *Time* cubic spline by *Group* and *Relatedness* [@wood2017generalized]. For both splines, we specified $k = 6$ basis functions or *knots*. @eq-model shows a formal implementation of the model.

$$
\begin{aligned}
\textbf{Likelihood:} \\
y_i &\sim \mathcal{N}(\mu_i, \sigma_i) \\ \\
\textbf{Linear model:} \\
\eta'(\mu_i) &= (\beta_0 + u _{0_{i}}) +
(\beta_1 + u _{1_{i}}) \cdot \text{Condition} +\\
&\beta_{2} \cdot \text{Group} +
\beta_{3} \cdot X + \beta_{4} \cdot (\text{Condition} \times \text{Group}) +\\
&\beta_{5} \cdot (\text{Condition} \times X) + \beta_{6} \cdot (\text{Group} \times X) +\\
&\beta_{7} \cdot (\text{Condition} \times \text{Group} \times X) +\\
& \sum_{j = 1}^k b_{j_{0}} \cdot \beta_{8_{k}} \cdot \text{Time} +
\sum_{j = 1}^k b_{j_{1}} \cdot\beta_{9_{k}} \cdot (\text{Time} \times \text{Group}) \\
\text{where:} \\
&\eta'\text{ is the empirical logit of target fixations} \\
&X \in \{\text{Age}, \text{L1 vocabulary}, \text{Total vocabulary}\} \\
&b_{j_{0-1}} \text{ are the basis cubic spline basis functions} \\
&k \text{ is the number of knots in the spline (10)} \\
\textbf{Prior:} \\
\beta_{0-9} &\sim \mathcal{N}(0, 0.5) \\
u_{0-2} &\sim \mathcal{N}(0, \sigma_{0-2})\\
b_{0-1} &\sim \text{MVN}(0, \tau_{0-1}) \\
\sigma_{0-2}, \tau_{0-1} &\sim \text{Exponential}(6)\\
\rho_{0-2} &\sim LKJCorr(6)\\
\text{where:}\\
&\rho_{0-2} \text{ are the correlation parameters for } \sigma_{0-2}
\end{aligned}
$$ {#eq-model}

The comparison between the three models including *Age*, *L1 vocabulary*, *Total vocabulary* as predictors was done using leave-one-out cross-validation (LOO-CV) as a benchmark of model performance, using Pareto-smoothed importance sampling (PSIS) to approximate it [@vehtari2017practical]. We implemented the model using `brms` [@burkner2017brms], an R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan]. We ran `r english::words(dim(model_fits_bcn$fit_0$fit)[2])` iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fits_bcn$fit_0), big.mark = ",")` iterations each and an additional `r format(niterations(model_fits_bcn$fit_0), big.mark = ",")` warm-up iterations per chain.

#### Statistical inference

We assessed the practical relevance of the estimated regression coefficients of the model following @kruschke2018bayesian. First, we specified a region of practical equivalence (ROPE) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`, in the probability scale. This region indicates the range of values that we considered equivalent to zero. We then summarised the posterior distribution of each regression coefficient with the 95% highest density interval (HDI). This interval contains the true value of this coefficient with 95% probability, given the data. Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, noted as $p(\text{ROPE})$, which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero). For example, $p(\text{ROPE})=.80$ indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero.

## Results

### Model comparisons

```{r tbl-loos}
#| label: tbl-loos
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of models 0 to 3 in Study 1."
get_loos <- function(x, y) {
	out <- lst(m_0 = x, m_1 = y) |> 
		loo_compare() |> 
		as.data.frame() |> 
		rownames_to_column("model") |>
		arrange(model) |> 
		select(model, elpd_diff, se_diff) |> 
		mutate(diff_label = sprintf("%.3f", elpd_diff),
			   se_label = sprintf("%.3f", se_diff))
	return(out)
}

loo_compare(model_loos_bcn[1:4]) |> 
	as.data.frame() |> 
	rownames_to_column("model") |> 
	select(model, matches("diff|elpd")) |> 
	mutate(model = paste0("Model ", gsub("fit_", "", model))) |>
	gt(rowname_col = "model") |> 
	fmt_number(is.numeric) |> 
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

@tbl-loos summarises the outputs of the LOO-CV model comparisons. Models $\mathcal{M}_{2}$ and $\mathcal{M}_{3}$, showed substantially better predicted performance than models $\mathcal{M}_{0}$ and $\mathcal{M}_{1}$, which indicates that adding the *Condition* predictor improves significantly the fit of the models. Both $\mathcal{M}_{2}$ and $\mathcal{M}_{3}$ models showed equivalent performance, suggesting that adding the two-way interaction between *Group*, and *Condition*, did not improve the the fit of the model substantially. We now report a description of the posterior distribution of the regression coefficients of model $\mathcal{M}_{3}$. We now report the median and 95% highest density interval of each fixed regression coefficient in $\mathcal{M}_{3}$ (see @tbl-coefs).


### Linear effects

```{r tbl-coefs}
#| label: tbl-coefs
#| tbl-cap: "Posterior distribution of regression coefficients. Median: median of the posterior distribution in logit scale. 95\\% HDI: 95\\% highest density interval of the distribution. *p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| out-width: 80%
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_lp1" = "Group",
				"b_condition1" = "Condition (U vs. RN)",
				"b_condition2" = "Condition (RN vs. RC)",
				"b_condition1:lp1" = "Group \U00D7 Condition (U vs. RN)",
				"b_condition2:lp1" = "Group \U00D7 Condition (RN vs. RC)")

model_summary <- get_posterior_summary(model_fits_bcn$fit_3,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval)

coefs <- model_summary |> 
	mutate(.value = sprintf("%.3f", .value),
		   .ci = sprintf("[%.3f, %.3f]", .lower, .upper),
		   .rope = scales::percent(.rope, accuracy = 0.01)) |> 
	split(janitor::make_clean_names(model_summary$.variable))

# summarise posterior draws
intercept_label <- paste0("Intercept (at ", round(mean(data_bcn$age), 2), " months)")

tbl_data <- model_summary |>
	select(.variable_name, .value, .lower, .upper, .rope, .type) |>
	mutate(.variable_name = gsub("×", "$\\\\times$", .variable_name),
		   .hdi = sprintf("[%.3f, %.3f]", .lower, .upper),
		   .rope = case_when(is.na(.rope) ~ "-",
		   				  .rope > 0.9999 ~ "1",
		   				  .rope < 0.00001 ~ "0",
		   				  .default = sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope)))) |>
	select(.variable_name, .value, .lower, .upper, .hdi, .rope)

col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")

tbl_data |>
	select(-c(.lower, .upper)) |>
	kable(
		format = "latex",
		escape = FALSE,
		col.names = col_names,
		digits = 3,
		booktabs = TRUE,
		align = "lclr",
		linesep = ""
	) |>
	kable_styling() |>
	add_footnote("$\\beta$: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution. $p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none", escape = FALSE)
```

Overall, the average participants' looking time exceeded chance levels, as indicated by the fact that the 95% HDI of the intercept term excluded zero ($\beta$ = `r coefs$b_intercept$.value`, 95% HDI = `r coefs$b_intercept$.ci`), and all of its posterior samples were outside of the ROPE. The coefficient of *Age* had a positive sign, but its 95% HDI overlapped completely with the ROPE ($\beta$ = `r coefs$b_age_std$mdn_label`, 95% HDI = `r coefs$b_age_std$.ci`), indicating that participants from all ages showed equivalent overall target word recognition. The 95% HDI of the coefficient of *Group* also included zero ($\beta$ = `r coefs$b_lp1$.value`, 95% HDI = `r coefs$b_lp1$.ci`) and completely overlapped with the ROPE, indicating an equivalent overall target preference in monolinguals and bilinguals. The 95% HDI of the first contrast of the *Condition* predictor---comparing *Unrelated* and *Related/Non-cognate* trials---included zero ($\beta$ = `r coefs$b_condition1$.value`, 95% HDI = `r coefs$b_condition1$.ci`), and `r coefs$b_condition1$.rope` of its posterior samples overlapped with the ROPE. The 95% HDI of the second contrast, comparing *Related/Non-cognate* and *Related/Cognate* trials, also included zero ($\beta$ = `r coefs$b_condition2$.value`, 95% HDI = `r coefs$b_condition2$.ci`), and `r coefs$b_condition2$.rope` of its posterior samples overlapped with the ROPE. The overall target preference was equivalent across both pairwise condition comparisons. The interaction term between the first *Condition* contrast contained zero ($\beta$ = `r coefs$b_condition1_lp1$.value`, 95% HDI = `r coefs$b_condition1_lp1$.ci`), with `r coefs$b_condition1_lp1$.rope` of its posterior samples overlapping with the ROPE. The interaction term between the second *Condition* contrast also contained zero ($\beta$ = `r coefs$b_condition2_lp1$.value`, 95% HDI = `r coefs$b_condition2_lp1$.ci`), and  `r coefs$b_condition1$.rope` of its posterior samples fell within the ROPE. The outcomes of this model provide strong to moderate evidence against differences in overall target looking time across conditions and across monolinguals and bilinguals.

A comparison between models including *Age* ($\mathcal{M}_{4}$), *L1 vocabulary* ($\mathcal{M}_{5}$), and *Total vocabulary* ($\mathcal{M}_{6}$)---and their three-way and two-way interactions with *Group* and *Condition*---against model $\mathcal{M}_{3}$, which only included *Age* as a covariate is shown in @tbl-loos-age. Overall, all models performed equivalently, with the simplest model $\mathcal{M}_{3}$ showing slightly better performance. This suggests that participants' target looking during the test phase can be predicted with relative accuracy without taking into account *L1 vocabulary*, or *Total vocabulary* sizes.

```{r tbl-loos-age}
#| label: tbl-loos-age
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of models 3 to 6 in Study 1."
loo_compare(model_loos_bcn[4:7]) |> 
	as.data.frame() |> 
	rownames_to_column("model") |> 
	select(model, matches("diff|elpd")) |> 
	mutate(model = paste0("Model ", gsub("fit_", "", model))) |>
	gt(rowname_col = "model") |> 
	fmt_number(is.numeric) |> 
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

Both younger and older participants showed equivalent looking time across conditions, as suggested by the fact that the 95% HDI of the two-way interaction between *Age* and the first *Condition* contrast included zero ($\beta$ = `r coefs$fit_4$b_condition1_age_std$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition1_age_std$ci_label`), as well as the interaction between *Age* and the second *Condition* contrast ($\beta$ = `r coefs$fit_4$b_condition2_age_std$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition2_age_std$ci_label`). This pattern of results was found in both monolinguals and bilinguals, as the 95% HDI of the three-way interaction between *Age*, *Group*, and *Condition* included zero both the first ($\beta$ = `r coefs$fit_4$b_condition1_lp1_age_std$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition1_lp1_age_std$ci_label`) and second ($\beta$ = `r coefs$fit_4$b_condition2_lp1_age_std$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition2_lp1_age_std$ci_label`) contrast of *Condition*.

### Time course of target looking

```{r timecourse}
make_std <- function(x, mean, sd) (x - mean) / sd

timebin_values <- make_std(seq(0, 11, length.out = 100),
						   mean(data_bcn$timebin),
						   sd(data_bcn$timebin))

age_values <- make_std(c(20, 30),
					   mean(data_bcn$age),
					   sd(data_bcn$age))

nd <- expand_grid(distinct(data_bcn, condition, lp),
				  age_std = 0, timebin_std = timebin_values)

epreds <- add_epred_draws(nd,
						  fit_3, 
						  re_formula = NA,
						  ndraws = 25)

# age_values <- make_std(c(20, 30),
# 					   mean(data_bcn$age),
# 					   sd(data_bcn$age))
# 
# nd <- expand_grid(distinct(data_bcn, condition, lp),
# 				  age_std = age_values,
# 				  timebin_std = timebin_values)
# 
# epreds <- add_epred_draws(nd, model_fits_bcn$fit_4, 
# 						  re_formula = NA,
# 						  ndraws = NULL,
# 						  value = ".elog")

epreds_diff <- epreds |> 
	pivot_wider(id_cols = c(lp, timebin_std,.draw),
				names_from = condition,
				values_from = .elog)

epreds_chance <- epreds |> 
	median_hdci() |> 
	mutate(is_lower_positive = .lower > 0) |> 
	filter(is_lower_positive) |> 
	summarise(over_chance = min(timebin_std),
			  .by = c(lp, condition, age_std)) |> 
	mutate(over_chance = (over_chance * 
						  	sd(data_bcn$timebin) +
						  	mean(data_bcn$timebin))*100+300) 

epreds_chance_report <- epreds_chance |> 
	mutate(over_chance = sprintf("%.2f", over_chance),
		   age_std = paste0("age_", gsub(" months", "", age_std)),
		   age_std = factor(x = age_std,
		   				 levels = age_values,
		   				 labels = paste0(c(20, 30), " months"))) |> 
	arrange(lp, age_std) %>%
	split(tolower(.$lp)) |> 
	map(\(x) split(x, x$age_std)) 
```

An analysis of the time course of target looking revealed a similar pattern of results (see @fig-epreds). Posterior mean prediction for the three conditions overlap across the full time course of the trial in both language groups, and also in both extrema of the age range of participants (20 and 32 months).

```{r fig-epreds}
#| label: fig-epreds
#| fig-cap: "Posterior mean predictions of the time course of target fixation in the test phase. Lines and intervals indicate the median and 95% credible interval of the posterior predictions, respectively."
#| fig-width: 8
#| fig-height: 4
epreds_summary <- epreds |> 
	ungroup() |> 
	summarise(.value = mean_qi(.epred)[[1]],
			  .lower = mean_qi(.epred)[[2]],
			  .upper = mean_qi(.epred)[[3]],
			  .by = c(lp, condition, timebin_std))

epreds |> 
	ggplot(aes(timebin_std, .epred,
			   colour = condition,
			   fill = condition,
			   shape = condition)) +
	facet_grid(~lp) +
	# geom_hline(yintercept = 0,
	# 		   linetype = "dashed") +
	# stat_summary(fun.data = mean_qi,
	# 			 geom = "ribbon",
	# 			 alpha = 1/4,
	# 			 linewidth = NA) 
	# geom_ribbon(aes(ymin = .lower,
	# 				ymax = .upper),
	# 			linewidth = NA,
	# 			alpha = 1/4) +
	geom_line(aes(group = interaction(.draw, condition)),
			  alpha = 1/5,
			  linewidth = 3/4) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 3/4) +
	# stat_summary(data = data_bcn,
	# 		 aes(y = plogis(.elog)),
	# 		 fun.data = "mean_se",
	# 		 geom = "errorbar",
	# 		 linewidth = 3/4,
	# 		 width = 0.1,
	# 		 position = position_dodge(0.1)) +
	stat_summary(data = filter(data_bcn, timebin_std < 0.5),
				 aes(y = .elog),
				 fun = "mean",
				 geom = "point",
				 size = 3)

stat_summary(fun.data = "mean_qi",
			 geom = "line")
stat_summary(aes(group = interaction(`child_id:session_id`,
									 condition)),
			 fun = mean,
			 geom = "line",
			 linewidth = 3/4,
			 alpha = 1/10) 
stat_summary(data = data_bcn,
			 fun = mean,
			 geom = "point",
			 size = 2.5) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition",
		 shape = "Condition") +
	# scale_x_continuous(breaks = make_std(seq(0, 17, 3),
	# 									 mean(data_bcn$timebin),
	# 									 sd(data_bcn$timebin)),
	# 				   labels = \(x) format(seq(3e2, 2e3, 3e2),
	# 				   					 big.mark = ",")) &
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  legend.position = "top")
```

In an exploratory analysis, we compared the model-estimated time points at which participants' gaze behaviour showed evidence of target looking above chance level (50%). We defined the recognition time of the trial as the earliest time point at which the lower bound of the 95% HDI posterior mean prediction of the target looking excluded zero. As expected, the recognition times were on average shorter for older infants than for younger infants. Differences between conditions were larger in bilinguals than in monolinguals, especially in the older group. The recognition time for monolingual 20-month-olds was `r epreds_chance_report$monolingual$age_20$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_20$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_20$over_chance[3]` ms in the *Related/Cognate* condition. In monolingual 30 month-olds, the recognition point was `r epreds_chance_report$monolingual$age_30$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_30$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_30$over_chance[3]` ms in the *Related/Cognate* condition. In bilingual 20 month-olds, the recognition point was `r epreds_chance_report$monolingual$age_20$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_20$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_20$over_chance[3]` ms in the *Related/Cognate* condition. In bilingual 30 month-olds, the recognition point was `r epreds_chance_report$monolingual$age_20$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_20$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_20$over_chance[3]` ms in the *Related/Cognate* condition.


```{r fig-recognition}
#| label: fig-recognition
#| fig-cap: "Model-predicted recognition times for monlingual and bilinguals at 20 and 30 months of age."
#| fig-width: 7
#| fig-height: 3
epreds_chance |> 
	mutate(age_std = factor(x = age_std,
							levels = age_values,
							labels = paste0(c(20, 30), " months"))) |> 
	ggplot(aes(lp, over_chance,
			   colour = sort(condition),
			   shape = sort(condition),
			   fill = sort(condition))) + 
	facet_wrap(~age_std) +
	geom_point(size = 3) +
	geom_line(aes(group = condition),
			  linewidth = 1,
			  alpha = 1/2) +
	labs(x = "Group",
		 y = "Recognition time (ms)",
		 colour = "Condition",
		 fill = "Condition",
		 shape = "Condition") +
	scale_y_continuous(breaks = seq(0, 1e3L, 50)) +
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  panel.grid.minor.y = element_line(colour = "grey",
		  								  linetype = "dotted"),
		  legend.position = "top",
		  axis.title.x = element_blank())
```

## Discussion

In Study 1, we adapted an implicit naming paradigm to investigate the potential role of within- and cross-language phonological priming in 21- to 32-month-old children. Both the overall target looking and its course time across the trial suggested strong word recognition, regardless of priming condition, participant language profile and age. Overall, participants' performance was equivalent across all priming conditions. An exploratory analyses on word recognition times suggested a complex pattern of results, showing an age effect---in which older participants looked at the target above chance earlier than younger infants---and a weak interaction between priming conditions, age, and language group. These results conflict with previous findings in several ways.

First, previous investigations using implicit naming paradigm had provided strong evidence in favour of infants and toddlers generating internal labels when presented with pictures of familiar objects. @mani2010infant and @mani2011phonological reported priming effects in 18 and 24-month-old infants. In primed trials, participants' target looking preference was influenced by the previous presentation of a phonologically related referent in silence. In control trials---in which prime and target labels were unrelated---no priming effect was found. This suggested that infants implicitly named the prime label, and that the resulting internal phonological form was detailed enough to interact with the subsequent auditory recognition of the target word. In the present study, we did not find evidence in favour of differences in target looking preference between related and unrelated trials, regardless of the age of participants or their language profile.

```{r doe-exposure}
doe_mon_exp <- participants_bcn |>
	filter(lp=="Monolingual")|> 
	unnest_wider(doe) |> 
	mutate(doe2 = ifelse(test_language=="Catalan",
						 doe_spanish, doe_catalan),
		   is_positive =  doe2 > 0) |> 
	count(is_positive) 
```

Both @mani2010infant and @mani2011phonological, and the present study share a similar design, but sampled different children populations. Participants in Mani and Plunkett's studies were raised in English speaking households in Oxford, while participants in the present study were learning Catalan or Spanish, if not a combination of both. We identify two ways in which this might have influenced the results. One is that the stimuli lists in Catalan and Spanish might have differed in some critical dimension, which interfered infants' ability to generate implicit labels for prime pictures. For example, words in the present study were predominantly bi-syllabic, while those from Mani and Plunkett's studies were mostly monosyllabic, and contained less phonemes on average. Longer word-forms might have been more difficult to retrieve in time for a prime effect to take place in target recognition. Another possibility is that exposure to a second language might have delayed the developmental trajectories of word production in bilinguals, in such way implicit naming emerges later ages. Given that `r scales::percent(doe_mon_exp$n[1] / doe_mon_exp$n[2], accuracy = 0.01)`
of monolinguals in Barcelona were also exposed to a second language, they might as well been affected by such delay in implicit naming trajectories.

The results from the Study 1 also conflict with previous reports of cross-language priming in toddlers. Particularly, @von2012language and @floccia2020translation had provided evidence of phonological priming through translation. One critical different between both studies and the present one is the silent or auditory presentation of the prime labels. In priming through translation studies, the prime is auditorily presented to participants, and cross-language priming effects take place through the subsequent activation of the prime translation. In the present study, the prime label was not named, and participants were expected to generate both the prime label and its translation implicitly. It may be the case that infants in the present study faced a more challenging task, having to internally retrieve phonological forms without any auditory prompt like the one provided by priming-through-translation studies. 

Finally, another possibility is that participants in the present study did in fact general prime labels, but such labels did not interfere over target recognition. This could be because they were not phonologically specified enough. This is unlikely, as previous studies have shown strong evidence of fine phonological representations in the early lexicon REFERENCE. Instead, it could be the case that the prime labels were activated, but phonological links between lexical representations are still weak or not present, so that priming effects are not possible until later in age. To our knowledge, this study is the first one to test bilingual children at this age on implicit naming paradigms, so evidence of priming is not present at the moment.

We found no differences between models including participants' age or any of the two indices of vocabulary size (L1 and total vocabulary size) as predictors. Previous studies suggest that same-language vocabulary size predicts better bilingual infants' performance in word recognition tasks, as compared to cross-language vocabulary size [@marchman2010vocabulary].

In summary, the possibly increased difficulty of the task induced by the properties of the stimuli, together with differences in task design and populations might have contributed to the null pattern of results shown in Study 1. Given the complexity of the picture drawn by these factors, we conducted a second study with the aim of replicating the original findings by @mani2010infant and @mani2011phonological. If phonological priming effects are observed in a same-aged population using the adaptation of the task from Study 1, this would provide evidence in favour of cross-language differences in the developmental trajectories of implicit naming. If, on the contrary, a similar patter of null results is found, this would suggest either the existence of finer differences in task implementation compared to the original studies, or point to the effect of interest being less robust than suggested by the literature. 

# Study 2

In this study, we conducted a conceptual replication of @mani2010infant and @mani2011phonological. Using the same task from Study 1, we tested a same-aged group of monolinguals being exclusively raised in British English speaking homes, the same population tested in the original studies. As in Study 1, we manipulated the phonological relationship between the prime and target labels at word onset, as well as the cognate status of the prime label. Evidence from implicit naming of prime images would be revealed by differences in target looking or its time course across the trials between *Unrelated* trials, and *Related/Non-cognate* and *Related/Cognate* trials. Given that participants in this study are monolinguals---and contrary to the monolinguals in Barcelona, have received no exposure to any other language---we would not expect the cognate status of the prime labels to exert any effect on subsequent target recognition.

## Methods

### Participants

```{r}
#| label: participants-numbers-oxf
participants_oxf <- filter(participants, location=="Oxford")

n_participants_total <- nrow(distinct(participants_oxf, child_id))
n_sessions_total <- count(participants_oxf)

n_participants_sex <- participants_oxf |> 
	distinct(sex, child_id) |> 
	count(sex)

n_participants_sessions <- participants_oxf |> 
	count(child_id, name = "n_sessions") |> 
	count(n_sessions) |> 
	group_split(n_sessions) |> 
	set_names(paste0("session_", 1:2)) |> 
	map("n")

ages <- participants_oxf |> 
	summarise(n = n(),
			  n_participants = n_distinct(child_id),
			  across(age, lst(mean, sd, min, max)),
			  across(matches("doe_"), lst(mean, sd)),
			  age_vctr = list(age),
			  .by = c(lp, test_language))

```

We collected data from `r n_participants_total` children (`r n_participants_sex$n[1]` female, `r n_participants_sex$n[2]` male, `r n_participants_sex$n[3]` sex not reported, age: *Mean* = `r round(ages$age_mean, 2)` months, *SD* = `r round(ages$age_sd, 2)`, *Range* = [`r round(ages$age_min, 2)`-`r round(ages$age_max, 2)`]), living in the Oxfordshire area (United Kingdom), tested at the Oxford BabyLab at the University of Oxford. Families were recruited from maternity rooms in private hospitals and social media, and contacted via phone when the child's age spanned between 20 and 32 months. From the `r n_participants_total` children that participated, `r n_participants_sessions$session_1` participated once, and `r n_participants_sessions$session_2` participated twice. Recurrent participants were tested with at least XXX months of difference. We gathered a total of `r n_sessions_total$n` testing sessions. All participants were being raised in exclusively British English monolingual homes.

### Vocabulary size

```{r vocab-values-oxf}
vocabulary_oxf <- inner_join(vocabulary, 
							 select(participants_oxf, session_id),
							 by = join_by(session_id))
```

We collected vocabulary data using parental responses to the Oxford Communicative Development Inventory (OCDI) [@hamilton2000infant]. Families were sent the questionnaire immediately after each experimental session, and were given two weeks to fill it. Since all vocabulary sizes of English monolingual infants were only assessed in English, both *L1 vocabulary* and *total vocabulary* sizes were identical (see @fig-vocabulary-oxf).

```{r fig-vocabulary-oxf}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Vocabulary size scores from the recurrent participants are shown linked."
#| fig-width: 4
#| fig-height: 4
vocabulary_oxf |> 
	inner_join(select(participants, child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |> 
	ggplot(aes(age_group, l1_prop)) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1/4, linewidth = 3/4) +
	geom_dots(dotsize = 2,
			  colour = clrs[5],
			  fill = clrs[5],
			  scale = 0.5,
			  side = "both",
			  alpha = 3/4) +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 linewidth = 3/4,
				 width = 0.1,
				 colour = "black",
				 show.legend = FALSE) +
	stat_summary(fun = mean,
				 geom = "point",
				 size = 2,
				 colour = "black",
				 show.legend = FALSE) +
	# geom_point(alpha = 1) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(labels = scales::percent) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```


### Stimuli

New stimuli lists were generated to adapt the experimental procedure to English-learning infants. Stimuli lists were created following the same constraints as in Study 1. Auditory stimuli were recorded following the same procedure as in Study 1, produced by Southern British English native female speaker in a toddler-directed manner. Audio and picture processing were performed following the same workflow as in Study 1.

### Procedure and apparatus

Same as in Study 1. The study was run using a custom Matlab script, PresentMate, based on the PsychToolbox-3 extension [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox]. Visual fixations were recorded using a Tobii TX 300 eye-tracker and a 23-in screen of 1920 $\times$ 1080 resolution. The Tobii Analytics SDK 3.0 was used to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 9-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented.

### Data analysis

Same as in Study 1, but removing the *Group* predictor from the models, as only monolingual infants were tested. 

```{r dataset_oxf}
attrition_oxf <- filter(attrition, location=="Oxford")
n_trials <- nrow(attrition_oxf)
n_sessions <- n_distinct(attrition_oxf$session_id)
n_participants <- n_distinct(attrition_oxf$child_id)
n_trials_valid <- nrow(filter(attrition_oxf,
							  is_valid_participant,
							  is_valid_trial))
n_sessions_valid <- attrition_oxf |> 
	distinct(child_id, .keep_all = TRUE) |> 
	pull(is_valid_participant) |> 
	sum()

n_participants_valid <- n_distinct(
	attrition_oxf[attrition_oxf$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition_oxf$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_oxf$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_oxf$is_valid_gaze_test_each)
n_exc_participants <- sum(!attrition_oxf |> 
						  	distinct(child_id, .keep_all = TRUE) |> 
						  	pull(is_valid_participant)) 

n_exc_participants_vocab <- attrition_oxf |> 
	distinct(child_id, .keep_all = TRUE) |> 
	mutate(is_valid_vocab_size = !is_valid_vocab_size) |> 
	pull(is_valid_vocab_size) |> 
	sum(na.rm = TRUE) 

n_longitudinal <- attrition_oxf |>
	filter(is_valid_participant,
		   is_valid_trial) |>
	distinct(child_id, session_n) |> 
	count(child_id, name = "times") |> 
	count(times)

n_trials_condition <- attrition_oxf |> 
	filter(is_valid_participant) |> 
	count(trial_type, is_valid_trial) %>%
	split(janitor::make_clean_names(.$trial_type))
```

We defined a time windows of interest from 200 ms after target and distractor pictures onset until the end of the trial at 2,000 ms when  both pictures disappeared from screen. To avoid modelling fixations driven by processes other than auditory word recognition [@fernald1998rapid; @fernald2001half]. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach], with a maximum of 20 maximum consecutive missing samples being interpolated (an equivalent of 166.67 ms).

We gathered data from `r format(n_trials, big.mark = ",")` trials from `r n_sessions` testing sessions, generated from `r n_participants` distinct participants. We excluded trials in which participants failed to provide 50% valid eye-tracking samples (equivalent to 750 ms) during the prime phase (*n* = `r format(n_exc_prime, big.mark = ",")`) or 50% valid samples (equivalent to 1,000 ms) during the target-distractor phase (*n* = `r format(n_exc_test, big.mark = ",")`). We also excluded trials in which participants did not provide at least 5% of valid samples (equivalent to 100 ms) of target or distractor looking in the test phase (*n* = `r format(n_exc_test_each, big.mark = ",")`) [see @floccia2020translation for a similar approach].

After trials that matched any of the aforementioned exclusion criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the  *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 10% (*n* = `r n_exc_participants_vocab`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final dataset dataset, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).

## Results

```{r results-oxf}
loos_tbl <- loo_compare(model_loos_oxf[1:2]) |> 
	as.data.frame() |> 
	rownames_to_column("model") |> 
	select(model, matches("diff|elpd"))
```

In line with the results in *Study 1*, including the *Condition* predictor increased the predictive performance of the model substantially, compared to a model including only the main effect of *Age* ($\text{ELPD}_{\mathcal{M}_1} - \text{ELPD}_{\mathcal{M}_0}$ = `r round(loos_tbl$elpd_diff[2], 2)`, *SE* = `r round(loos_tbl$se_diff[2], 2)`). We now report the median and 95% highest density interval of each fixed regression coefficient in $\mathcal{M}_{3}$.

```{r tbl-coefs-oxf}
#| label: tbl-coefs-oxf
#| tbl-cap: "Posterior distribution of regression coefficients. Median: median of the posterior distribution in logit scale. 95\\% HDI: 95\\% highest density interval of the distribution. *p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| out-width: 80%
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_condition1" = "Condition (U vs. RN)",
				"b_condition2" = "Condition (RN vs. RC)")

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_1,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval)

coefs <- model_summary |> 
	mutate(.value = sprintf("%.3f", .value),
		   .ci = sprintf("[%.3f, %.3f]", .lower, .upper),
		   .rope = scales::percent(.rope, accuracy = 0.01)) |> 
	split(janitor::make_clean_names(model_summary$.variable))

# summarise posterior draws
intercept_label <- paste0("Intercept (at ", round(mean(data_oxf$age), 2), " months)")

tbl_data <- model_summary |>
	select(.variable_name, .value, .lower, .upper, .rope, .type) |>
	mutate(.variable_name = gsub("×", "$\\\\times$", .variable_name),
		   .hdi = sprintf("[%.3f, %.3f]", .lower, .upper),
		   .rope = case_when(is.na(.rope) ~ "-",
		   				  .rope > 0.9999 ~ "1",
		   				  .rope < 0.00001 ~ "0",
		   				  .default = sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope)))) |>
	select(.variable_name, .value, .lower, .upper, .hdi, .rope)

col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")

tbl_data |>
	select(-c(.lower, .upper)) |>
	kable(
		format = "latex",
		escape = FALSE,
		col.names = col_names,
		digits = 3,
		booktabs = TRUE,
		align = "lclr",
		linesep = ""
	) |>
	kable_styling() |>
	add_footnote("$\\beta$: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution. $p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none", escape = FALSE)
```

Overall, the average participant' looking time exceeded chance levels, as indicated by the fact that the 95% HDI of the intercept term excluded zero ($\beta$ = `r coefs$b_intercept$.value`, 95% HDI = `r coefs$b_intercept$.ci`) and all of its posterior samples fell outside of the ROPE. The 95% HDI of the coefficient of *Age* had a positive sign, but did not exclude zero ($\beta$ = `r coefs$b_age_std$.value`, 95% HDI = `r coefs$b_age_std$.ci`), and overlapped completely with the ROPE, indicating that participants from all ages showed equivalent overall target word recognition. The 95% HDI of the first contrast of the *Condition* predictor---comparing *Unrelated* and *Related/Non-cognate* trials---included zero ($\beta$ = `r coefs$b_condition1$.value`, 95% HDI = `r coefs$b_condition1$.ci`), and `r coefs$b_condition1$.rope` of its posterior samples fell within the ROPE. The 95% HDI of the second contrast---comparing *Related/Non-cognate* and *Related/Cognate* trials---also included zero ($\beta$ = `r coefs$b_condition2$.value`, 95% HDI = `r coefs$b_condition2$.ci`), and `r coefs$b_condition2$.rope` of its posterior samples fell within the ROPE. The overall target preference was equivalent across both pairwise condition comparisons.

```{r tbl-loos-oxf-age}
#| label: loos-oxf-age
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of the models in Study 2."
loo_compare(model_loos_oxf[2:4]) |> 
	as.data.frame() |> 
	rownames_to_column("model") |> 
	select(model, matches("diff|elpd")) |> 
	mutate(model = paste0("Model ", gsub("fit_oxf_", "", model))) |>
	gt(rowname_col = "model") |> 
	fmt_number(is.numeric) |> 
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

@loos-oxf-age shows a comparison between models including *Age* ($\mathcal{M}_{2}$), and *L1 vocabulary* ($\mathcal{M}_{3}$)---and their two-way and two-way interaction with *Condition*---against model $\mathcal{M}_{1}$, which only included *Age* as a covariate suggested that all models performed equivalently, with model $\mathcal{M}_{2}$ showing a slightly better performance. Overall, this suggests that participants' target looking during the test phase can be predicted with relative accuracy without taking into account *L1 vocabulary*, or *Total vocabulary* sizes.


```{r timecourse-oxf}
timebin_values <- make_std(seq(0, 17, length.out = 100),
						   mean(data_oxf$timebin),
						   sd(data_oxf$timebin))

nd <- expand_grid(distinct(data_oxf, condition),
				  timebin_std = timebin_values,
				  age_std = 0)

epreds <- add_epred_draws(nd, model_fits_oxf$fit_oxf_1, 
						  re_formula = NA,
						  ndraws = NULL,
						  value = ".elog")

epreds_chance <- epreds |> 
	median_hdci() |> 
	mutate(is_lower_positive = .lower > 0) |> 
	filter(is_lower_positive) |> 
	summarise(over_chance = min(timebin_std),
			  .by = c(condition, age_std)) |> 
	mutate(over_chance = (over_chance * 
						  	sd(data_oxf$timebin) +
						  	mean(data_oxf$timebin))*100+300) 

epreds_chance_report <- epreds_chance |> 
	mutate(over_chance = sprintf("%.2f", over_chance),
		   age_std = paste0("age_", 
		   				 gsub(" months", "", age_std))) |> 
	arrange(age_std) %>%
	split(tolower(.$age_std)) 
```

An analysis of the time course of target looking revealed a similar pattern of results (see @fig-epreds). Posterior mean prediction for the three conditions overlap across the full time course of the trial in both language groups, and also in both extrema of the age range of participants (20 and 32 months).

```{r fig-epreds-oxf}
#| label: fig-epreds-oxf
#| fig-cap: "Posterior mean prdictions of the time course of target fixation in the test phase. Lines and intervals indicate the median and 95% credible interval of the posterior predictions, respectively."
#| fig-width: 8
#| fig-height: 4
epreds |> 
	ggplot(aes(timebin_std, .elog,
			   colour = condition,
			   fill = condition)) +
	geom_hline(yintercept = 0, linetype = "dashed") +
	stat_summary(fun.data = mean_qi,
				 geom = "ribbon",
				 alpha = 1/4,
				 linewidth = NA) +
	stat_summary(fun = mean,
				 geom = "line",
				 linewidth = 3/4) +
	stat_summary(data = data_oxf,
				 fun = mean,
				 geom = "point") +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = make_std(seq(0, 17, 3),
										 mean(data_bcn$timebin),
										 sd(data_bcn$timebin)),
					   labels = \(x) format(seq(3e2, 2e3, 3e2),
					   					 big.mark = ",")) &
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  legend.position = "top")
```

In an exploratory analysis, we compared the model-estimated time points at which participants' gaze behaviour showed evidence of target looking above chance level (50%). We defined the recognition time of the trial as the earliest time point at which the lower bound of the 95% HDI posterior mean prediction of the target looking excluded zero. As expected, the recognition times were on average shorter for older infants than for younger infants. Differences between conditions were larger in bilinguals than in monolinguals, especially in the older group. The recognition time for monolingual 20-month-olds was `r epreds_chance_report$monolingual$age_20$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_20$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_20$over_chance[3]` ms in the *Related/Cognate* condition. In monolingual 30 month-olds, the recognition point was `r epreds_chance_report$monolingual$age_30$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_30$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_30$over_chance[3]` ms in the *Related/Cognate* condition. In bilingual 20 month-olds, the recognition point was `r epreds_chance_report$monolingual$age_20$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_20$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_20$over_chance[3]` ms in the *Related/Cognate* condition. In bilingual 30 month-olds, the recognition point was `r epreds_chance_report$monolingual$age_20$over_chance[1]` ms in the *Unrelated* condition, `r epreds_chance_report$monolingual$age_20$over_chance[2]` ms in the *Related/Non-cognate* condition, and `r epreds_chance_report$monolingual$age_20$over_chance[3]` ms in the *Related/Cognate* condition.


```{r fig-recognition-oxf}
#| label: fig-recognition-oxf
#| fig-cap: "Model-predicted recognition times for monlingual and bilinguals at 20 and 30 months of age."
#| fig-width: 4
#| fig-height: 3
epreds_chance |> 
	ggplot(aes(age_std, over_chance,
			   colour = sort(condition),
			   fill = sort(condition))) + 
	geom_point(shape = 16, size = 4) +
	geom_line(aes(group = condition),
			  linewidth = 1) +
	# geom_point(size = 3) +
	labs(x = "Group",
		 y = "Recognition time (ms)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_y_continuous(breaks = seq(0, 1e3L, 25)) +
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  panel.grid.minor.y = element_line(colour = "grey",
		  								  linetype = "dotted"),
		  legend.position = "top",
		  axis.title.x = element_blank())
```

## Discussion

Study 2 was aimed at investigating whether the phonological priming effects previously reported @mani2010infant and @mani2011phonological could be found using the adaptation of the implicit naming task form Study 1 in a sample with equivalent demographic characteristics to the original studies. Results from Study 2 did not provide any evidence in favour of phonological priming. English monolingual participants from all ages showed an equivalent pattern of target looking across both related and unrelated trials.

The absence of a phonological priming effect suggest that either English monolinguals did not generate implicit labels for the prime pictures presented in silence, or that if generated, such labels did not interact with the subsequent recognition of the target word. Both candidate explanations conflict with both Mani and Plunkett's studies, and also with previous studies suggesting that infants 12-months and older already generate internal labels when presented with pictures of familiar objects.

One of the main differences between the implicit naming paradigm used by @mani2010infant and the one in Studies 1 and 2 is the fact that target labels were auditorily presented before target and distractor pictures were presented on screen. Given that priming effects are very sensitive to the timing of the activation of the lexical presentation [@REFERENCES], we decided to shorten the time span between prime image naming and target label presentation by removing the pre-naming interval of the test phase, in which both target and distractor pictures are presented form a baseline time interval before the target label is auditorily presented. By presenting the auditory target label immediately after the offset of the prime picture, the internally generated prime phonological form would be more likely to affect the activation of the target lexical representation. It is possible that this difference in task design might have led to the contrasting outcomes of Mani and Plunkett and the present study. Given that participants were not presented with the target and distractor pictures until after the target label was auditorily presented, one possibility is that participants faced the additional task of exploring both pictures to find the target referent, which infants in @mani2010infant did not have to do, because of the pre-naming presentation of target and distractor in the test phase. This visual exploration of the screen might have increased the cognitive load of the task, rendering the potential effect of phonological priming to a weaker, undetectable effect.

Overall, the results from Study 2 support a more methodological explanation of the null pattern of results in Study 1, in which the lack of a phonological priming effect---both within and across languages---might be explained by differences in task design or replicability concerns.

# General discussion

In this paper, we investigated the developmental trajectories of cross-language co-activation in the developing lexicon. We tested a longitudinal an cross-sectional sample of 20-to-32 month olds in an implicit naming paradigm, in which we designed three trial types to manipulate the phonological overlap between the prime and target words within and across languages. In *Unrelated* trials, prime and target were phonologically unrelated in both languages. In *Related/Non-cognate* trials, prime and target labels shared phonological onset only in the dominant language of participants, in which they were tested. In *Related/Cognate* trials, the prime label was a cognate: prime and target labels shared phonological onset in both languages. In Study 1, we tested a cohort oh monolingual and bilingual infants learning Catalan, Spanish, or both, and found no phonological priming effect in monolinguals or bilinguals. In Study 2, we tested a cohort of English monolinguals, and found similar results. 
To our knowledge this is the first study investigating cross-language priming using an implicit naming paradigm. Such paradigm presents a critical asset for studying bilingual lexical processing. Since the prime pictures are presented in silence, the task is exclusively complete in the dominant language of participants. This provides a monolingual situation in which differences in performance across experimental conditions that involve cross-linguistic interactions can be more easily associated to the experimental manipulation in a causal way [@grosjean1997bilingual]. This contrast to other experimental paradigms in which participants are presented with---need to produce---items from both languages. Such covert manipulation of the similarity between stimuli from the two languages has allowed previous studies to provide strong support for the language-non selective account of lexical access e.g., [@costa2000cognate; @thierry2007brain; @von2014bilinguals].  

Previous studies have provided evidence of English monolingual infants 18-month-old and older being sensitive to phonological priming effects during auditory word recognition. @mani2010infant had shown a facilitation effect in word recognition when 18-month-olds were previously presented with a familiar picture whose label shared phonological onset with the target word. @mani2011phonological extended these results to 21-month-olds, showing an effect in the opposite direction: phonologically related primes interfered with target word recognition. The shift from facilitation to inhibition was interpreted as a shift from sub-lexical facilitation at the phonological level---in which target word recognition is facilitated by the prior activation of an overlapping set of phonological segments---to a lexical interference effect in which the activation of the lexical representation of the phonologically similar prime led to competition dynamics with the subsequently activated representation of the target word. The last conclusion was supported by the fact that the size of the interference effect was associated with the increase in participants' vocabulary size, and also with the cohort size of the prime label---primes belonging to a larger neighbourhood of phonologically related words were subject to stronger interference effects during the task

Even earlier evidence of implicit naming was provided through ERP registering at 14 months of age. @duta2012erp presented name-known pictures in silence for one second. After one second, the picture's label was presented auditorily, and the infants' N400 component was registered. In some trials the auditory label corresponded to a canonical realisation of the word-form. In other trials, the word was mispronounced. Infants ERP signal differed when presented with correct pronunciations or mispronunciations, indicating that infants had generated expectations at the phonological level about the auditory labels of the picture before they were presented with its auditory label. A later study by @styles2015infant re-analysed the first 1,000 ms of the trials in the previous study (while the picture was being presented before the auditory label was presented), and found that the ERPs of name-known pictures (according to parental reports of receptive vocabulary) differed from those of name-unknown pictures (which infants would not be able to lexicalise). This provides further support to the account that at around 14 months of age, infants already name familiar pictures, even if presented in silence. In bilinguals, only one study had used an implicit naming paradigm to study cross-language priming. @von2014bilinguals adapted @mani2010infant procedure to  test German-English bilinguals. The authors manipulated the phonological overlap between the prime label in English (presented in silence) and a German target (auditorily presented) in both German and English. Participants' ERPs revealed differences between trials in which the German prime label overlapped phonologically with the English target label, compared to trials in which none of the labels overlapped. These results suggest that adult bilinguals implicitly named prime pictures, and that the resulting label interacted with subsequent word recognition.

The results in the present investigation are in strong contrast with the aforementioned studies. Both studies 1 and 2 conflict with implicit naming. 


Study 1 conflicts with cross-language studies. singh2014,

@kuipers2015bilingualism compared ERP responses to picture-word pairs in English monolinguals and Welsh-English bilinguals where the words were either semantically related or unrelated to a pictured object. Although both studies tested 30-month-olds, the conclusions the authors reached do not fully agree. @singh2014one concluded that there is a discontinuity between adults and young children in the way the two lexicons interact, while @kuipers2010event and @kuipers2015bilingualism concluded that there are no differences in adults’ and toddlers’ lexical-semantic processing.

The present study presents methodological contributions. First, the longitudinal design of the data collection provided a more suitable design for drawing conclusions about developmental change in word recognition trajectories, compared to studies that tested participants from a single age group [@floccia2020translation], or studies that with an exclusively cross-sectional design [@von2012language]. Second, the modelling approach in the present study incorporates the double source of repeated measures simultaneously into the random effects structure of the multilevel model. The models in Studies 1 and 2 accounts for the time course of target looking across the trials and its associated autocorrelation using Hierarchical General Additive Models [@barr2008analyzing; @pedersen2019hierarchical]. The models also incorporate the observations from the same testing session as a nested grouping variable inside participants, so that the test session-level parameters in each model are estimated taking into account the estimations of other testing sessions from the same participants. Overall, this approach allowed us to simultaneously model the participant-level variability, the testing session-level variability, and the time series consisting of participants' gaze behaviour during the task. Finally, the adoption of a Bayesian approach to implement and estimate the model allowed the incorporation of prior knowledge about the distribution of the parameters to generate stable estimates, despite the complexity of the model.


Why. Auditory presentation, implicit label generation but no priming

rhyme vs onset

word length

Another possibility is that the current study lacked the appropriate statistical power to detect the effect sizes of interest. Reported differences in looking time associated to priming paradigms in the available literature in children are generally small. Despite the relatively larger sample size achieved in the present investigation, participant-level trajectories of target looking are highly variable, which might have increased the uncertainty of the models' estimates.

The findings in the present investigation have important implications for current models of bilingual lexical development. For instance, recent accounts of bilingual words acquisition pointed to a facilitative role of phonological similarity between the two languages. A monograph by @floccia2018vocabulary provided evidence of a language distance effect, where bilinguals learning two languages sharing higher lexical similarity knew more words in their less dominant language than those learning two language with lower lexical similarity. Later work by @mitchell2022cognate, @siow2022effect, and @garcia-castro2023cognate provided evidence for a more direct facilitation effect of cognateness. Cognates are form-similar translation equivalents. Cognateness is suggested to facilitate word acquisition through parallel activation: the co-activation of translation equivalents in speech might lead to an faster accumulation of learning instances with words, which ultimately results in an earlier age of acquisition. In the present study, we failed to observed cross-language effects during word recognition. These outcomes are not in line with parallel activation taking place during speech exposure in bilinguals. It is possible that the mechanisms involved in the cognate facilitation in bilingual word acquisition are not captured by the task used in the present study. Implicit naming might be tapping into lower-level functions, more related to acoustic and phonological processing, while word acquisition and learning might rely on more abstract sources of information that do recruit parallel activation. 


recent work using parental reports of word acquisition have provided insights.  For instance, previous work on lexical acquisition has provided some insight into how  found that English-Spanish bilinguals' vocabulary in English was influenced by the words they had already acquired in Spanish. This is consistent with previous accounts of lexical acquisition in which infants acquire new words through preferential acquisition [@hills2009longitudinal, @fourtassi2020growth]. 


One of the limitations of this study lies in the use of cognates as prime words. 




# Conclusions




# References




