```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggdist)
library(patchwork)
library(here)
library(readr)
library(dplyr)

options(readr.show_progress = FALSE,
		readr.show_col_types = FALSE)

model_summary <- read_csv(here("data", "02-chapter-2",
							   "results", "posterior", "model_summary.csv"))

posterior_doe_summary <- read_csv(here("data", "02-chapter-2",
									   "results", "posterior", "posterior_summary_doe.csv"))

posterior_doe_draws <- read_csv(here("data", "02-chapter-2",
									 "results", "posterior", "posterior_draws_doe.csv"))


my_theme <- theme_minimal() +
	theme(panel.grid = element_line(colour = "grey",
									linetype = "dotted"),
		  axis.line = element_line(colour = "black"),
		  text = element_text(size = 12, colour = "black"),
		  axis.text = element_text(colour = "black"))

theme_set(my_theme)

```

As a robustness check, we fit a model similar to the one described in the main manuscript, but including lexical frequency and language degree of exposure as separate predictors, instead of the composite measure *Exposure*. Language degree of exposure (*DoE*) was included in interaction with *Age* and *Cognateness*, while lexical frequency (*Frequency*) was included as a main effect. @tbl-coefs-doe shows a comparison between the posterior distribution of the regression coefficients of both models. Overall, results are equivalent. 
\newpage

\blandscape
\begingroup
\small
```{r tbl-coefs-doe}
#| label: tbl-coefs-doe
#| tbl-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately. $\\beta$: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution. $p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
tbl_data <- list("Exposure model" = model_summary,
				 "DoE-Frequency model" = posterior_doe_summary) |> 
	purrr::map_df(\(x) mutate(x, across(.variable_name, as.character)),
				  .id = ".model") |> 
	mutate(across(c(.value, .lower, .upper),
				  function(x) {
				  	ifelse(.type=="Intercepts (at 22 months)",
				  		   plogis(x),
				  		   x/4)
				  })) |>
	filter(.type=="Slopes") |> 
	select(.model, .variable_name, .value, .lower, .upper, .rope) |> 
	mutate(across(.value:.upper, plogis),
		   .variable_name = gsub("×", '$\\\\times$', .variable_name),
		   .hdi = paste0("[", round(.lower, 3),
		   			  ", ", round(.upper, 3), "]"),
		   .rope = sub("^(-?)0.", "\\1.",
		   			sprintf("%.3f", .rope))) |> 
	select(.variable_name, .value, .lower, .upper, .hdi, .rope) 

col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")

tbl_data |> 
	select(-c(.lower, .upper)) |> 
	knitr::kable(format = "latex",
				 escape = FALSE,
				 col.names = col_names,
				 digits = 3,
				 booktabs = TRUE,
				 align = "lclr",
				 linesep = "") |> 
	kableExtra::pack_rows("Model: Exposure",
						  start_row = 1,
						  end_row = 8) |> 
	kableExtra::pack_rows("Model: Frequency & DoE",
						  start_row = 9,
						  end_row = 17) 
```
\endgroup
\elandscape