```{r setup}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

participants <- read_csv(here("data", "03-chapter-3", "data", "participants.csv"))
vocabulary <- read_csv(here("data", "03-chapter-3", "data", "vocabulary.csv")) |> 
	mutate(across(child_id, as.character))
attrition_trials <- read_csv(here("data", "03-chapter-3", "data", "attrition_trials.csv"))
attrition_participants <- read_csv(here("data", "03-chapter-3", "data", "attrition_participants.csv"))
bvq_data <- readRDS(here("data", "03-chapter-3", "data-raw", "bvq.rds"))
```

Given the large number of participants in Study 2 whose families failed to fill in the vocabulary questionnaire in time, we imputed missing vocabulary sizes to avoid losing their data in subsequent analyses. We used the `mice` R package [@buuren2011mice] to perform single imputations via Bayesian linear regression. @fig-vocabulary-imputation shows the distribution of observed and imputed vocabulary sizes. Overall, the distribution of imputed vocabulary sizes was equivalent to the distribution of observed vocabulary sizes of participants with similar age and language profile.

```{r fig-vocabulary-imputation}
#| label: fig-vocabulary-imputation
#| fig-cap: "Imputation of missing receptive vocabulary size scores in Study 2 for monolinguals (A) and bilinguals (B). vocabulary size scores are epxressed in raw counts, that is, the number of words reported by caregivers as acquired. Vocabulary sizes were imputed from a larger sample of children who provided responses to the Barcelona Vocabulary Questionnaire. We used participants' age and language profile to impute scores of participants wih the same profile. Imputed observations are highlighted in red (monolinguals) and blue (bilinguals)"
#| fig-height: 6
#| fig-width: 6
vocabulary_tmp <- vocabulary |>
	inner_join(select(participants, session_id, vocab_id, lp, age),
			   by = join_by(session_id, vocab_id)) |>
	relocate(session_id) |>
	filter(is_imputed)

clrs <- c("#004AAD", "#C8102E", "#FF9E1F")

fig_data <- bvq_data$vocabulary |> 
	select(-contents) |> 
	inner_join(distinct(bvq_data$logs, child_id, response_id, lp, age),
			   by = join_by(response_id)) |> 
	mutate(is_imputed = FALSE) |> 
	bind_rows(vocabulary_tmp) |> 
	select(-matches("prop")) |> 
	pivot_longer(ends_with("_count"),
				 names_to = "measure",
				 values_to = "vocab") |> 
	mutate(is_imputed = ifelse(is_imputed, "Imputed", "Observed"),
		   measure = factor(measure,
		   				 levels = c("total_count",
		   				 		   "l1_count"),
		   				 labels = c("Total",
		   				 		   "L1"))) |> 
	drop_na(vocab, measure) |>
	filter(between(age, 10, 40),
		   lp != "Other") |> 
	select(age, lp ,measure, vocab, is_imputed) |> 
	arrange(desc(is_imputed)) 

plot_mon <- fig_data |> 
	filter(lp=="Monolingual") |> 
	ggplot(aes(age, vocab, 
			   colour = is_imputed,
			   fill = is_imputed)) +
	facet_wrap(~measure, 
			   scales = "free_y") +
	geom_point(size = 2,
			   shape = 1,
			   stroke = 1)  +
	scale_colour_manual(values = c(clrs[2], "grey90")) +
	theme(axis.title.x = element_blank())

plot_bil <- fig_data |> 
	filter(lp=="Bilingual") |> 
	ggplot(aes(age, vocab, 
			   colour = is_imputed,
			   fill = is_imputed)) +
	facet_wrap(~measure, 
			   scales = "free_y") +
	geom_point(size = 2,
			   shape = 1,
			   stroke = 1)  +
	scale_colour_manual(values = c(clrs[1], "grey90")) +
	theme(strip.text = element_blank())

(plot_mon / plot_bil) +
	plot_layout(ncol = 1,
				guides = "collect") &
	plot_annotation(tag_levels = "A") &
	scale_fill_manual(values = c(clrs[2], "grey90")) &
	scale_x_continuous(breaks = seq(0, 50, 5)) &
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 fill = "Imputed",
		 colour = "Imputed") &
	theme(legend.position = "none",
		  legend.title = element_blank(),
		  panel.grid.minor = element_blank())

```